### 2020-12-20 src/paper.tex
initial commit

% sage_latex_guidelines.tex V1.20, 14 January 2017

\documentclass[Afour,sagev,times]{sagej}

\usepackage{moreverb,url}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

% -- to allow highlight changed sections in yellow:
\usepackage{color,soul}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2020}

%%Vancouver (numbered)
\bibliographystyle{SageV}
%\bibliography{paper.bib}

\begin{document}

\runninghead{Charlton and Laudan}

\title{A Web-Based Data Visualization Platform for MATSim}

\author{Billy Charlton\affilnum{1} and Janek Laudan\affilnum{1}}

\affiliation{\affilnum{1}Technische Universit\"at Berlin}

\corrauth{Billy Charlton, Technische Universit\"at Berlin,
Transport Systems Planning and Transport Telematics,
Secretariat SG 12,
Salzufer 17-19, 10587 Berlin, Germany}

\email{charlton@tu-berlin.de}

\begin{abstract}
There are many tools available for analyzing MATSim transport
simulation results, both open-source and commercial. This research
builds a new open-source visualization platform for MATSim outputs that
is entirely web-based. After initial experiments with many different web
technologies, a client/server platform design emerges which leverages
the advanced user interface capabilities of modern browsers on the
front-end, and relies on back-end server processing for more processor-intensive
tasks. The initial platform is now operational and includes several
aggregate-level visualizations including origin/destination flows,
transit supply, and emissions levels; as well as a fully disaggregate
traffic animation visualization. These visualizations are general
enough to be useful for various projects. Further work is needed
to make them more compelling and the platform more useful for practitioners.
\end{abstract}

\keywords{MATSim, data visualization, agent-based simulation}

\maketitle

\section{Introduction}

MATSim\cite{R1} is an open-source framework for implementing large-scale agent-based transport simulations. MATSim is widely used for transportation research in academic settings, and is gaining momentum as a tool ready for practice in real-world planning contexts.

There are many tools available for analyzing MATSim results, both
open-source and commercial. Typically, analysts can choose either
the free tool OTFVis\cite{R2} or the commercial software
Via \cite{R3}, both of which are desktop software packages requiring installation as well as a fair amount of technical acumen to operate. Alternatives to these tools include the more general-purpose desktop mapping software packages such as QGis and ArcGIS, or statistical software packages, again all of which require installation and expertise to use.

As MATSim moves from the confines of academic research to a more
public-facing role, a notable gap is apparent: there are no web-based interactive tools available for disseminating MATSim data and results. This creates a challenge for using MATSim in public policy settings: the only people who can meaningfully examine and explore results are those who have extensive technical knowledge and access to the specialized software and large datasets involved.

This research explores one way to fill this gap: building an open
web-based data visualization platform which is specifically designed
to complement MATSim.

\section{Motivation}

The rapid advancement of Internet browsing technologies in the last
five years has enabled the web browser to do things much more
"application"-like than ever before: background processing,
three-dimensional rendering using \hl{graphics card (GPU)} acceleration, offline
support, and more. The combination of these technologies and their
standard implementations on every popular hardware type and operating system now makes the web a very compelling platform.

For MATSim research, the question is: could a web browser really
be useful for exploring and delivering results when the datasets
are so large? Answering this question is the primary motivation for
this research. Essentially, has the web become powerful enough for MATSim?

Currently, analysis of MATSim outputs ends up in research reports, \hl{Adobe (PDF)} documents, video screen-recordings, and presentations. An online dashboard of results that a user could explore and manipulate would not only be more interactive, but might also reveal findings that the original analysts hadn't anticipated.

\section{Requirements}
The research team at Technische Universit\"at Berlin had several
"blank slate" discussions before any code was written: meaning,
if we could start at the very beginning and design something completely web-based and open, what would the bare minimum requirements be for it to be truly useful? The following requirements emerged from those discussions.

\subsection{Requirement 1: Web browser-based}

Given the above-stated motivation and hypothesis that the modern web
platform is ready for large-scale visualization tasks, the most obvious requirement is that the product of this research must work with any modern web browser.

Several specific web technologies developed and made widely available in recent years enable us to perform this research: HTML 5, CSS 3, WebGL, ECMA Script 6, and Web Workers. Briefly, these technologies are:

\begin{enumerate}

\item[(i)] \textbf{HTML 5} improves and standardizes the "document model" of what constitutes a web page and how it is specified.
\item[(i)] \textbf{CSS 3} is a styling language that enables fine-grained layout and styling of individual elements on a page. CSS 3 defines in a consistent, standard way the details of elements including color, size, layout, and animation of page elements.
\item[(ii)] \textbf{WebGL} provides in-browser support for the 3D-accelerated graphics capabilities of modern machines.
\item[(iv)] \textbf{ECMAScript 6} is an updated (2015) version of the venerable Javascript scripting language that has been part of the web platform since the early 1990's. Recent versions of Javascript eliminate the more problematic aspects of the language and make it easier for developers to create bug-free, efficient code.
\item[(v)] \textbf{Web Workers} are a recent (2013) addition to the web platform that allow background processing for long-running tasks. Before Web Workers, there was no way to run truly multi-threaded code inside a browser.

\end{enumerate}

A complication in web development is that the major web browser vendors implement these technologies on their own timelines, some much more rapidly than others. Further complicating things is the reality that end users do not always upgrade their browsers frequently (or at all). This creates a landscape where there is a technology adoption curve with a very long "tail". Thus, developers of every web site need to make a conscious decision about where to
draw the line -- choosing necessary technologies for their site to operate correctly, while knowing that some users with older browsers will either have a sub-optimal experience or no access to the site at all.

For this research, we are deliberately exploring the \textit{latest standard} web technologies, with the expectation that access to these technologies will become more and more common in the future. Thus, we are targeting the most recent versions of modern web browsers as of 2019, including Google Chrome, Mozilla Firefox,
and Apple Safari. All three browsers fully support the above-listed technologies, and importantly, all three auto-update automatically, ensuring that most users of those browsers stay current as these technologies evolve. \hl{This does leave out older browsers such as Internet Explorer and Edge, as well as browsers with small market share. Supporting those browsers would require much more deliberate and time-consuming testing, which while not impossible, is currently out of scope. For some government users this is acknowledged as possibly problematic, since use of Internet Explorer 11 is still mandated at some agencies.}

\subsection{Requirement 2: Open source}

The entire project, including all front-end (browser) and back-end (server) code, must be fully open-source.

No proprietary or closed licensing schemes were considered, primarily because excellent proprietary visualization packages for MATSim already exist. Creating a competing product would be duplicative and unnecessary, and would not further the research goal of determining whether web-based technology is now advanced enough to work with MATSim outputs. The goal of this research is not to replace existing proprietary solutions, but rather to complement them.

The software developed as part of this research is licensed entirely with the GNU General Public License version 3, commonly known as the \textit{GPL v3}\cite{R4}.

This matches the license of MATSim itself. Several other open-source licenses were considered, including the MIT License and the Apache Public License, but the benefit of sharing a common license with MATSim outweighs any perceived benefits of switching to other open licenses \hl{The main benefit of those other open licenses is compatibility with commercial software. Given its narrow scope, this research is not likely to be of interest to existing commercial vendors. Furthermore, the research team does not intend to turn this into a commercial product. Thus the GPL license is a good fit.}

\subsection{Requirement 3: Use good defaults, with minimal configuration, and be opinionated}

Since its inception, the web platform has had a relentless focus on simplicity and smooth user onboarding. Users are accustomed to being immediately familiar with a site -- often within seconds of their first interaction. Because of this expectation, it is critical that this research follow current best practices
for user interface (UI) and user experience (UX). Specifically, that means using familiar UI paradigms such as navigation bars and breadcrumbs, separating configuration from usage, limiting settings and options to the bare minimum, and being "opinionated", i.e. encouraging a correct way to accomplish a task.

This approach is dissimilar to some data exploration tools (e.g., QGis and Via) where extreme configurability is emphasized. Rather than providing myriad options for details such as scales and color ramps, our research focuses on choosing good defaults and determining whether that is sufficient for the software to be
useful.

\subsection{Requirement 4: An extensible platform}

Every data visualization use case is different; there is no way to anticipate how the tool will be used. If the platform is too generic, it will be not at all useful. Conversely, if only hard-coded visualizations are created for specific projects, it will be relegated to "demo-ware", meaning it is a successful technology demonstration but not actually useful for real users.

To fulfill this requirement the software platform will need to be extensible: basic capabilities and templates will be provided, but a user with some level of coding skill should be able to create new visualizations that are not anticipated by the researchers.

\section{Initial experiments}

It is no exaggeration to state that the Javascript code library ecosystem is extremely, enormously large. Thousands of libraries and packages are available on a common centralized Javascript repository known as "NPM", and there are often multiple packages that do similar things. As a developer, one must assess and select from these packages or choose to solve a problem by writing code by hand. Of course these libraries are of varying levels of popularity and quality.

Based on the requirements laid out above, some initial experiments were carried out to assess various approaches before committing to a technology stack.

\subsection{Visualizing time-dependent data on a geographic map}

Two popular web-based Javascript libraries were tested for displaying geographic data; Leaflet and Mapbox GL. A simple test case comprised of MATSim simulation outputs was developed, with the goal of displaying aggregated roadway link volumes by time of day.

Leaflet (\url{leafletjs.com}) is very popular and its application programming interface (API) is a bit simpler than that of Mapbox GL. Leaflet uses background map "tiles" at specific zoom levels, and layers data on top of those base maps. With small networks (we tested Cottbus, Germany, a small city of 100,000 inhabitants) Leaflet performed well, but medium-sized and large-sized networks with many elements visible at once suffered from noticeable performance degradation. This was problematic, as this was the simplest use-case envisioned.

Mapbox GL (\url{mapbox.com}) fared much better, apparently better-suited to displaying large datasets with many visible features simultaneously. In addition, Mapbox GL's use of 3D vector graphic mapping instead of preset tiles made for a much more pleasing user experience, with smooth animations between zoom levels and better background processing during page loads. For these reasons, Mapbox GL was chosen as the base map for the remaining geographic visualizations.

\subsection{Visualizing non-geographic data}

There are hundreds of data visualization libraries available for the web which provide ways to produce charts and plots of varying complexity. Our requirement of using open-source code narrows the field considerably.

After experimenting with several packages including D3, Raphael, Morris and others, the package Vega-Lite\cite{R5} exhibited many of the characteristics desired. Notably, Vega-Lite follows a declarative syntax known as the "grammar of graphics", as popularized by Wilkinson\cite{R6}. This grammar allows concise description of the meaningful components of a graphic. An added benefit is that graphs and charts can be downloaded in image format or in Vega source format, which is helpful for other researchers wanting to learn how to use the data format.

\subsection{Dealing with large datasets}

MATSim network files are small enough to fit in \hl{computer memory (RAM)}, but MATSim plan files and event files can be much larger than RAM, necessitating careful consideration about how to handle them.

Modern browsers allow access via API to a data storage area that is unique per hostname, i.e. http://mysite.com is allowed some storage on the local machine. Initial experiments revealed that each browser vendor implements this storage differently, with very strict limits on the absolute amount storage available, sometimes based on how much free space remains on the user's machine. It became apparent that this local browser-based storage would not be sufficient for MATSim outputs. Running a local file-server process would allow browser access to a specific folder on the machine, which might be nice for advanced analysts but violates the research goal of being fully web-based on the client end. Thus, a client-server paradigm emerges as the only truly viable alternative, and indeed this is how most websites operate today: the browser is the front-end to the heavier processing and storage tasks that happen on someone else's server. Note that "someone else's server" is usually referred to as "the cloud".

\section{Platform architecture}

A client/server architecture was chosen for this research based on the initial experiments described above.

The research team authored back-end server software for file storage, user authentication, and data pre-processing. Due to space considerations, this paper does not delve into the details of those components. Suffice it to mention that the front-end communicates with them to establish what resources a user has access to, and provides an application programming interface, or API, with which to query and fetch available files and resources. The code for those servers is also open source, and may be the subject of future papers.

The front-end architecture has several interacting components, described below.

\subsection{"Vue" Single Page Application}

The primary framework used to build the application is known as "VueJS" or Vue (available at \url{vuejs.org}). Vue is a framework for building interactive user interfaces on the web; it provides the glue between the items a user clicks and the code that runs when they do so. Vue provides many services which allow a web page to behave more like a full-featured application, including state management, routing between different URLs, and componentization of code in a way that encourages code reuse and loose coupling. Vue depends on Javascript, which means it \hl{does not work for users} who have disabled scripting in their browser.

Vue is most often used to write so-called "single page applications" which are websites that behave like desktop applications. Most large, popular websites such as Github, Twitter, and Facebook all employ this paradigm, meaning the site handles page transitions and URLs as if they are all in one common namespace, and the user thinks less about visiting URLs and more about navigation through visual components to accomplish tasks. This matches our use case.

Vue components each encapsulate the three elements required for the modern web: HTML layout, Javascript code, and CSS formatting. Components only interact with each other through well-defined pathways of properties and events, which greatly improves debuggability.

\subsection{Build system}

The build system of a modern web application is fairly complex and the Javascript ecosystem changes rapidly. After numerous iterations, the current build system comprises a series of individual tools that all work together to produce the final assets that get sent to a user's browser. Those tools include the Vue command line interface (CLI), the NPM package manager, webpack, babel, and TypeScript.

Notably, the initial codebase was migrated to the Typescript language midway through development, as the benefits of a strongly-typed language were perceived to be worth the development time. TypeScript is a separate language from JavaScript, and enforces type annotations for variables and adds additional features such as enumerations. The TypeScript compiler then converts TypeScript code into ECMAScript 6 JavaScript, which can be run in the browser (as JavaScript is the only scripting language that browsers support).

\subsection{Visualization plug-ins}

One of the main requirements of this research is to produce a system where new visualizations can be produced rapidly and added into the existing framework to generate new capabilities.

The Vue component architecture enables this. To create a new visualization, a developer copies an existing "blank" visualization template and gives it a new name, specifies the file inputs and parameters required, and then uses the above-described libraries to modify the code per their needs.

This currently requires ample coding skill in Javascript; it is not a system that is point-and-click like an online data exploration tool. Experience with other similarly-designed platforms suggests that software-minded analysts or modelers would be able to extend the platform, but typical end users will not.

\section{Results: the current state of the tool}

A working instance of the platform is now online and available at \url{viz.vsp.tu-berlin.de}. Sample datasets are uploaded, and pre-built visualizations are publicly accessible, as a demonstration of the platform's current state. There is also a user login system so that internal researchers can extend and experiment with the system, without exposing data or work-in-progress to the public.

Basic user, project, and file management capabilities are operational. This includes grouping files by model run or by other user-defined tags, as can be seen in Figure~\ref{F1}.

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-1.png}
\caption{Project dashboard}\label{F1}
\end{figure}

See the following screenshots in Figures~\ref{Fod}-~\ref{Fanim} for examples of the current state of the user interface. Note of course that the tool demos better live than in screenshots, as the whole point of this research is to develop an interactive tool.

Several types of aggregate visualizations are developed:

\begin{enumerate}

\item[(i)] Origin/destination flows between aggregate areas, so-called "spider diagrams" (Figure~\ref{Fod})
\item[(i)] Link flows, by time of day and mode (Figure~\ref{Flinkvols})
\item[(iii)] Transit supply explorer, which displays all transit routes and allows the user to see which routes serve specific stops and links. (Figure~\ref{Ftransit})
\item[(iv)] Sankey diagrams, which can be used to depict changes/flows between between scenarios across multiple choices, such as shifts in mode between two scenarios (Figure~\ref{Fsankey})
\item[(v)] Emissions levels on a geographic hexagonal grid basis (Figure~\ref{Femissions})

\end{enumerate}

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-2.png}
\caption{Origin/Destination flows}\label{Fod}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-3.png}
\caption{Link volume summary}\label{Flinkvols}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-4.png}
\caption{Transit depiction}\label{Ftransit}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-5.png}
\caption{Mode shift diagram}\label{Fsankey}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-6.png}
\caption{Emissions plot}\label{Femissions}
\end{figure}

In addition, one disaggregate animation is available:

\begin{enumerate}

\item[(i)] A vehicle flow simulation, showing individual vehicle agents in real-time on the network. (Figure~\ref{Fanim})

\end{enumerate}

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-7.png}
\caption{Vehicle animation}\label{Fanim}
\end{figure}

\subsection{Workflow feedback}

User interface design (UI) and user experience (UX) have their own entire fields of research. Initial prototypes of the tool simply did not meet the needs of test users. Problems included: difficult discerning which files were which; inability to efficiently use the model run tagging feature (in which a user could mark sets of files as belonging to a particular MATSim run); separate user logins causing data to be inaccesible to team members working on the same projects (resulting in everyone using a common "team" login, against best practices); private projects "leaking" onto the public website; and myriad usability bugs in the data visualizations themselves.

These usability problems were eventually solved by surveying other technical websites which organize and present large amounts of data. One website in particular, \url{Github.com}, was found to be well-liked by test users and has a similar hierarchical view of data: users can belong to organizations, and both organizations and users can create projects ("repositories" on Github) which may hold large numbers of files.

By adopting a project and user paradigm similar to that employed by Github, users were immediately more familiar with the site and had less to learn. The current site adheres to this so closely that test users often refer to the platform as "MatHub".

\subsection{Performance}

Even with modern hardware and the latest browsers, it is quite challenging to produce performant, visually pleasing results with disaggregate MATSim data. The vehicle flow simulation depends heavily on the back-end server to produce and deliver simulation "frames" to the browser in real-time, so that the browser simply has to render the data.

Various levels of aggregation make MATsim data much easier to visualize, as is reflected in the number of different visualizations this research was able to produce with aggregate data in a short time frame.

\textbf{Preprocessing.} The traffic animation and the emissions "hex grid" visualizations both rely on back-end server processes to preprocess the raw MATSim event outputs. This takes anywhere from a few seconds to many minutes, depending on the size of the simulation that was run. The preprocessing only needs to occur once, and thereafter the results are cached and stored on the server. The UI presents a helpful "still processing" message during this stage. Unfortunately, changing some settings such as the size of the hex-grids for the emission tool means the calculations must be redone. Further work is being done to make these processes run as quickly as possible.

\textbf{Map layer limitations.} The Mapbox GL mapping library allows lines, points, and polygons to be displayed on top of a base map. There seems to be a number of map elements beyond which performance becomes very slow; various techniques can be used such as aggregating elements at further-out zoom levels to get around this. Another option is to use Mapbox GL only for the base map, and to use the WebGL graphics libraries to draw large numbers of elements directly on top of the map. This is the technique used for the traffic animation, which can easily support tens of thousands of elements (all in motion!) simultaneously. For additional visualizations which have large numbers of grahical elements, more research will need to be done on layering WebGL elements on top of the base map.

\subsection{Mobile device support}

Initially, the research targeted desktop browsers only. The reasoning was that analysts currently use desktop software, and it would be sufficient to complement those desktop tools with a new web-based option.

However, as the tool started becoming usable by internal users, it quickly became apparent that everyone wanted some version of the visualizations to work on mobile devices, too. It was suggested that even a read-only visualization "viewer" for mobile would be better than not having the tool work at all on mobile.

To address this, the layout of the main user interface needed to be redesigned, but the overall stack of web-based libraries and components chosen was already well-suited to mobile use. This effort is currently underway.

\section{Conclusions and outlook}

Experimenting with the various technologies and getting all of the disparate pieces working together was an enormous task, one which took much longer than anticipated. However, those decisions are now behind us and the platform has become quite stable.

A new visualization can now be generated by the researchers in a matter of days or weeks. The researchers are admittedly very familiar with the inner workings of the system, but even so it has been encouraging to see new visualizations go from ideation to rough draft in such a short time.

None of the above-listed visualizations are particularly groundbreaking or visually stunning. And, all of them could be easily created in other tools (although usually without the interactivity that the web enables). This is a bit disappointing but the open nature of the platform, requiring no software installation by end users, still has an advantage: it opens up the display of MATSim results to the public and to decisionmakers, even if they do not have access to desktop mapping or travel forecasting software.

Another use case that has emerged from these visualization experiments is a more regimented data management tool for MATSim. Currently there is no straightforward way to share MATSim datasets online. The combination of the new file storage and management capabilities along with the Github-like user interface provides a natural place for users to store and disseminate results.

Importantly, the world has not stood still while this platform was under development. Just in the past year, major data visualization efforts from well-funded companies such as Uber and others have been released. There are legitimate questions about how much of this work could be superceded by large, well-funded, professional coding teams.

Despite these concerns, the MATSim visualization framework is operational and is now just beginning to be useful for researchers at the department of its creation. This bodes well for further development in the near future.

All code is available on the MATSim Github site, at \url{github.com/matsim-org/viz}.

\subsection{Author Contribution Statement}

The authors confirm contribution to the paper as follows: study conception and
design: Billy Charlton, Janek Laudan; data collection: Billy Charlton; analysis and
interpretation of results: Billy Charlton, Janek Laudan; draft manuscript
preparation: Billy Charlton. All authors reviewed the results and approved
the final version of the manuscript.

% -------------------------------------------------------------

\begin{acks}
This research project is part of the National Research Programme "Big Data" (NRP 75) of the Swiss National Science Foundation.\\
Website: \url{http://www.nrp75.ch/}
\end{acks}

% \bibliography{paper}
%% \bibliographystyle{SageV}

%-----------------------------
% \iffalse
\begin{thebibliography}{1}

\bibitem{R1}
Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016). \textit{The multi-agent transport simulation MATSim} (p. 618). London: Ubiquity Press.

\bibitem{R2}
Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 225-–234.

\bibitem{R3}
Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–-224.

\bibitem{R4}
Free Software Soundation (2007) \textit{GNU General Public License}, Version 3.
URL: www.gnu.org/licenses/gpl.html

\bibitem{R5}
Satyanarayan, A., Moritz, D., Wongsuphasawat, K., and Heer, J. (2016) \textit{Vega-Lite: A Grammar of Interactive Graphics}. IEEE Transactions on Visualization and Computer Graphics, Volume 23, Issue 1. DOI: doi.org/10.1109/TVCG.2016.2599030

\bibitem{R6}
Wilkinson, Leland (2005) \textit{The Grammar of Graphics}, Second Edition. Springer Press, Chicago, USA. DOI: doi.org/10.1007/0-387-28695-0

\end{thebibliography}
%\fi
%----------------------------

\end{document}

### 2020-12-29 src/ChargingFreight.tex
bibtex is working

% Template for Elsevier CRC journal article
% version 1.2 dated 09 May 2011

% This file (c) 2009-2011 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Computer Science

% Changes since version 1.1
% - added "procedia" option compliant with ecrc.sty version 1.2a
%   (makes the layout approximately the same as the Word CRC template)
% - added example for generating copyright line in abstract

%-----------------------------------------------------------------------------------

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          %%
%% Important note on usage                                  %%
%% -----------------------                                  %%
%% This file should normally be compiled with PDFLaTeX      %%
%% Using standard LaTeX should work but may produce clashes %%
%%                                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
%% The 'procedia' option causes ecrc to approximate to the Word template
\documentclass[3p,times,procedia]{elsarticle}
\flushbottom

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
\usepackage[bookmarks=false]{hyperref}
    \hypersetup{colorlinks,
      linkcolor=blue,
      citecolor=blue,
      urlcolor=blue}
%\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[utf8]{inputenc}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Procedia Computer Science}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{W. Charlton}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
%\jnltitlelogo{Computer Science}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols

\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{authoryear}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}
%\usepackage{harvard}
% put your own definitions here:x
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}
\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{FIXME The 9th International Workshop on Agent-based Mobility, Traffic and Transportation Models, \\ (ABMTRANS) March 23 - 26, 2021, Warsaw, Poland}%
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
%% \dochead can also be used to include a conference title, if directed by the editors
%% e.g. \dochead{17th International Conference on Dynamical Processes in Excited States of Solids}

\title{Open-Source Web-Based Visualizer for Dynamic-Response Shared Taxi Simulations}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[a]{William Charlton}

\address[a]{Technische Universität Berlin, Chair of Transport Systems Planning and Transport Telematics, Straße des 17. Juni 135, 10623 Berlin, Germany}

%% ----- ABSTRACT -----
\begin{abstract}
We developed a unique, web-based public outreach tool used by transit agencies investigating future shared-taxi fleet scenarios. The tool is in operation for several cities and rural regions including Berlin and several small regions in western Germany. The project goal is to explore the possibility of augmenting or even replacing fixed-route transit lines in less-developed areas with automated, connected, shared taxi fleets. Since those fleets don’t yet exist outside of Arizona, our team used an agent-based transport microsimulation model (MATSim) to develop various scenarios and produce results.
\end{abstract}

\begin{keyword}
shared taxis; data visualization; matsim; public outreach

%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

%\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000.}
\cortext[cor1]{Corresponding author. Tel.: +49 30 314-23308 ; fax: +49 30 314-26269.}
\email{charlton@vsp.tu-berlin.de}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text

%\enlargethispage{-7mm}
\section{Introduction}
\label{introduction}

\section{Methodology}
\label{methodology}
It is apparent that the charging strategy plays a major role in answering the posed research question. Therefore, this research focuses on designing the charging infrastructure for the respective MATSim scenario by \citet{MartinsturnerEtAl2020ETrucksFoodABMTrans} by applying and analyzing multiple different approaches to the charging of trucks.
MATSim is an extensible, activity- and multi-agent based transport simulation, which enables the simulation of large scale scenarios. It follows an iterative process several times until a stable state is reached. Synthetic agents are modeled which have previously defined schedules for their daily activities. The executed plans result from the agent's choice of transport mode and from interactions with other agents (e.g. traffic jam). After the simulation period (one day), the agents’ plans are analyzed and scored. In the next step, the agents can replan their daily schedule by changing the transport mode or route \cite{MATSimBook}. MATSim covers i.a. passenger transport \cite{ZiemkeEtAl2019OpenBerlinScenario} and freight transport \cite{SchroederLiedtke2014FoodDistributionBerlin, ZilskeJoubert2015FreightTrafficInBook}.

%\begin{figure}[ht]\vspace*{4pt}
%%\centerline{\includegraphics{fx1}\hspace*{5mm}\includegraphics{fx1}}
%\centerline{\includegraphics[width=0.8\textwidth]{Lademethodik_Schaubild_SW3.png}}
%\caption{Process diagram of the charging methodology}
%\label{fig:lademethodik}
%\end{figure}

A two-step method is developed (Figure \ref{fig:lademethodik}). First, the activity types and locations, the driven distances, and the number of vehicles, which are at the same time at the same location are derived from the respective MATSim scenario. We use current market data from series production and prototypes to define the battery capacity, the possible charging power, the consumption and the range of the vehicles. The charging demands result from the previously defined vehicle consumption and their simulated mileages from MATSim \cite{Jahn.2020}. Thereafter, we differentiate public charging and depot charging. For public charging, an algorithm identifies the position and number of charging stations by locally resolving the charging demands from MATSim (presented in \cite{Jahn.2020}). For depot charging, the location of the depot defines the position of the charging stations. The number of charging stations has to be defined afterwards. Commercial fleets usually perform charging in depots. The maximum number of charging stations is equal to the number of vehicles if each vehicle has its own charging station for overnight charging. Additionally, the installation of fast charging points enables charging during short-standing times in between tours. The number of vehicles that are simultaneously at the depot and require charging at daytime determines the number of fast charging points. The fast charging points can also be used at nighttime. The maximum number of normal charging stations can be obtained by subtracting the fast charging points from the number of vehicles at this depot (assumption: every vehicle has its own charging station). This number can be minimized by using each charger sequentially and charging several vehicles one after the other. For this purpose, the time that can be used for charging (vehicles are in the depot) is multiplied by the charging power of all chargers at the respective depot to derive the maximum possible charging capacity. Now the charging demands of all vehicles are analyzed and the vehicles are assigned to the charging points. While regarding maximum charging power and available time, the vehicles are charged in a way that the charging capacity is used sufficiently without exceeding it. This problem corresponds to the “subset sum problem”, where a certain number of items should be selected from a set of items in order to reach a target value as high as possible without exceeding it \cite{ALBERTOCAPRARA}. The algorithm passes through the list of trucks and keeps all trucks in the list whose total consumption is less or equal to the possible charging capacity. In this way, a few trucks are distributed to the first charging station. The remaining trucks are distributed iteratively to the other charging stations. If the minimum number of charging points necessary for the sequential overnight charging is less or equal to the number of fast chargers for the opportunity charging in the day time, this is the final number of charging points. If not all vehicles can be charged on the fast chargers sequentially, slow chargers for all remaining vehicles are added. The total number of chargers is the sum of the number of chargers in all depots.
As the cost of charging stations rises with the offered charging power, the definition of the minimal number and power of stations needed is of high economic value. Therefore, the state of charge (SOC) and the remaining tour lengths of the vehicles are analyzed. Thus, it is possible to determine the minimum energy demand for each vehicle arriving at the depot. Since the time required to pick up new goods is known, the required changing power for the  vehicle can be determined. Since charging stations are available at certain power levels, we run through different scenarios with different charging power levels and check how many of the vehicles can be charged in this way.

\section{Case Study}
\label{casestudy}
\citet{MartinsturnerEtAl2020ETrucksFoodABMTrans} include different vehicle types for which we define specific properties such as the battery capacity, and the consumption (see  Table \ref{tab:specs}). To investigate the effects of high power charging, the maximum charging power for all vehicles is set to 1,000 kW. This might not reflect the manufacturer's data of the reference vehicles. However, it reflects the current state of the art and research for heavy BEVs \cite{Zhu.21720202202020}. First, we analyze the transport simulation in terms of activity duration and charging demands. We define that charging infrastructure is only available at the depots of the carriers. For this reason, we consider the standing time of the trucks in the depots as the possible time span for (re)charging (Figure \ref{fig:duration}). Next, we quantify the energy consumption of the trucks before they reach the depot.  Depending on the traffic situation, a specific consumption is considered in the calculations. The maximum speed allowed on the link is compared to the simulated average speed. This ratio allows a statement about the traffic situation. The consumption is adjusted accordingly. If the average speed is much lower than the maximum speed, this indicates stop and go traffic, and increased consumption by 30 \% on this link is assumed. A similar conclusion was reached by \citet{Li.2014}, who analyzed stop and go traffic and its effects. The consumption distribution of the different vehicles is shown in Figure \ref{fig:consumption}. The optimum charging power is determined by analyzing what percentage of vehicles with a certain charging power can manage their route. For complete electrification, we assume that the trucks are able to charge at high performance opportunity chargers (HPOC) at strategic points in the city (access to city highways etc) if they are not able to complete their routes otherwise. The number of chargers and their charging power is determined by using the above-mentioned method for public charging.
\begin{table}[ht]
\caption{Vehicle specifications \cite{MartinsturnerEtAl2020ETrucksFoodABMTrans}}
\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}lllll@{}}
\toprule
Vehicle type & Light Duty & Medium Duty & Medium Duty & Heavy Duty\\
\colrule
Weight [t] &  7.5   &   18  &  26  &  40\\
Battery capacity gross (net) [kWh]  &  87 (60.9)   &   122 (85.4)  &  286 (200.2)  &  443 (310.1)\\
Energy consumption [kWh/m] &  0.00061   &   0.00106  &  0.00150  & 0.00180\\
\botrule
\label{tab:specs}
\end{tabular*}
\end{table}

%\begin{figure}[ht]\vspace*{4pt}
%\centerline{\includegraphics{fx1}\hspace*{5mm}\includegraphics{fx1}}
%\centerline{\includegraphics[width=0.7\textwidth]{images/duration_pickup.png}}
%\caption{Deviation of pickup duration}
%\label{fig:duration}
%\end{figure}

%\begin{figure}[ht]\vspace*{4pt}
%\centerline{\includegraphics{fx1}\hspace*{5mm}\includegraphics{fx1}}
%\centerline{\includegraphics[width=0.7\textwidth]{images/consumption.png}}
%\caption{Deviation of consumption of electric trucks}
%\label{fig:consumption}
%\end{figure}

%\begin{figure}[ht]
%\centering
%\begin{subfigure}[c]{0.49\textwidth}
%    \includegraphics[width=\linewidth]{images/duration_pickup.png}
%    \caption{Distribution of pickup duration for reloading goods}
%    \label{fig:duration}
%\end{subfigure}
%\begin{subfigure}[c]{0.5\textwidth}
%    \includegraphics[width=\linewidth]{images/consumption.png}
%    \caption{Distribution of consumption of electric trucks}
%    \label{fig:consumption}
%\end{subfigure}
%\caption{Distribution of pickup duration and consumption of electric trucks}
%\end{figure}

\section{Results}
\label{results}
After tracking the SOC of the trucks, it can be seen that the SOC of some trucks still drops below 0 \% which indicates that not all trucks can complete their route (Fig. \ref{fig:soc}). Therefore, it is not possible to charge all vehicles at the depot even with high charging power as shown in Figure \ref{fig:percentage}. For this reason, this number of vehicles is not considered for planning the infrastructure at the depot. Furthermore, it shows that even with a relatively low charging power of 200 kW or 400 kW a high percentage of vehicles can be sufficiently charged. Charging powers of over 600 kW do not provide better performance of the fleet. As some trucks are below a SOC 0 \%, the calculation of the number of charging stations is executed once with all trucks (279 trucks) and once only with the trucks that make their route (248 trucks at 400 kW charging power). The number of trucks depends on the charging power and can be seen in Figure \ref{fig:percentage}. 248 trucks require a maximum of 181 normal chargers and 61 fast chargers or a minimum of 67 fast chargers, see Table \ref{tab:final_amount}. By installing additional charging points, the entire truck fleet (all 279 trucks) in Berlin can cover its planned route. The number of additional charging stations is related to the charging power of the depot charger and is displayed in Table \ref{tab:additional_amount}. The results show that the supply of food retailing stores using electric vehicles is possible in urban areas. Over 90~\% of conventional trucks can be replaced by electrified trucks when recharging during the day at the depots is applied.
We show that 67 fast charging points with 400 kW charging power would be needed in the depots. The analysis shows that with a good occupancy rate, about four trucks can use a single charging point overnight. Even with a higher charging power of over 600 kW, the number of charging stations cannot be further reduced because it is limited by the charging stations that are needed simultaneously during the day. The difference of 6 chargers is the result of several depots that do not need fast charging during the day (because the tours starting from these depots are rather short) but still need chargers for over night charging. Charging stations with a power of over 400kW  would therefore be over-dimensioned. Chung \cite{Chung.20200928} comes to a similar conclusion in his analysis.
Since the case study of the food retailing in Berlin contains some really long routes, state of the art BEV-technology cannot complete these tours, even with intermediate charging at the depot. This explains the permanent deviation in Figure \ref{fig:percentage}. By setting up a further 31 HPOCs in public places, all electrified trucks could manage their route. 400 kW charging power is also sufficient for these stations if a 30-minute charging stop is planned. However, this would mean that the trucks that have an extra charging stop would end their tours up to 30 minutes later. Other than the charging during loading times, these extra 30 minutes would be unproductive time.

%\begin{figure}[ht]
%\centering
%\begin{subfigure}[c]{0.5\textwidth}
%    \includegraphics[width=\linewidth]{images/soc.png}
%    \caption{SOC of all trucks over the day with 400 kW charging power}
%    \label{fig:soc}
%\end{subfigure}
%\begin{subfigure}[c]{0.49\textwidth}
%    \includegraphics[width=\linewidth]{images/can_charged_new.png}
%    \caption{Percentage of vehicles that can be charged with specific charging power}
%    \label{fig:percentage}
%\end{subfigure}
%\caption{SOC analysis of all trucks with depot charging only}
%\end{figure}

\begin{table}[ht]
\caption{Resulting number of chargers for the different charging strategies}
\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}llllll@{}}
\toprule
 & Trucks  & Normal charger & Fast charger & HPOC\\
\colrule
Maximum (used by only one vehicle overnight) &  248   &  187  &  61  &  -\\
Minimum (400 kW charger used by multiple vehicles overnight)  &  248   &  -  &  67  &  -\\
Maximum with additional charger (used by only one vehicle overnight) &  279   &  214  &  65  &  31\\
Minimum with additional charger (400 kW charger used by multiple vehicles overnight)  &  279   &  -  &  71  &  31\\
\botrule
\label{tab:final_amount}
\end{tabular*}
\end{table}

\begin{table}[ht]
\caption{Additional HPOCs depending on the charging power}
\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}lllll@{}}
\toprule
Charging power depot charger [kW] & 200 & 400 & 600 & 800 \\
Number of HPOCs & 36 & 31 & 27 & 27 \\
\botrule
\label{tab:additional_amount}
\end{tabular*}
\end{table}

\section{Conclusion and Outlook}
\label{conclusion}
We presented a methodology to determine possible charging strategies for urban freight transport. Therefore, we provided current literature on charging and transport simulation. We analyzed the MATSim case study on the food retailing industry and found that with 31 HPOCs all tours can be operated. In the depots, 214 normal combined with 65 fast charging points are sufficient to charge the trucks. If the vehicles share the charging points overnight, this number is reduced to 71 fast charging points with 400 kW. With higher charging power, the share of fully charged vehicles hardly increases.
Our results show that the electrification of the food trade is technically feasible. However, it is necessary to consider the economic side, which was not taken into account in this analysis. This could be investigated in detail with a Total Cost of Ownership analysis. For even better performance, the next step is to consider the ranges of the truck types in the MATSim model using the range constraint presented in \cite{EwertEtAl2020_VRPDistanceConstraint_accepted} in combination with the here presented method to determine adequate charging strategies. It is expected that the number of trucks and thus also the number of charging stations will increase slightly, since the 9 \% of trucks that would currently not be able to complete their route, will be distributed among several trucks.This way however, the unproductive recharging time at HPOCs could be avoided. Further research should investigate the limits of electrification, e.g. by running scenarios with BEVs with a higher battery capacity, even if this results in less payload due the higher weight of the batteries. This could be extended by running the electrification scenarios with electric trucks with different battery capacities for each vehicle type. The presented method is transferable to various transport sectors. MATSim scenarios including for example personal commercial transport could be analyzed as well.

\section*{Acknowledgement}
This work was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) -- 398051144.

%\bibliography{bibliography_freight.bib}
\bibliographystyle{elsarticle-harv}

\begin{thebibliography}{1}

  \bibitem{R1}
  Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016). \textit{The multi-agent transport simulation MATSim} (p. 618). London: Ubiquity Press.

  \bibitem{R2}
  Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 225-–234.

  \bibitem{R3}
  Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–-224.

  \bibitem{R4}
  Free Software Soundation (2007) \textit{GNU General Public License}, Version 3.
  URL: www.gnu.org/licenses/gpl.html

  \bibitem{R5}
  Satyanarayan, A., Moritz, D., Wongsuphasawat, K., and Heer, J. (2016) \textit{Vega-Lite: A Grammar of Interactive Graphics}. IEEE Transactions on Visualization and Computer Graphics, Volume 23, Issue 1. DOI: doi.org/10.1109/TVCG.2016.2599030

  \bibitem{R6}
  Wilkinson, Leland (2005) \textit{The Grammar of Graphics}, Second Edition. Springer Press, Chicago, USA. DOI: doi.org/10.1007/0-387-28695-0

  \end{thebibliography}

\end{document}

### 2020-12-29 src/procs-template-ABMTRANS_2021.tex
bibtex is working

% Template for Elsevier CRC journal article
% version 1.2 dated 09 May 2011

% This file (c) 2009-2011 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Computer Science

% Changes since version 1.1
% - added "procedia" option compliant with ecrc.sty version 1.2a
%   (makes the layout approximately the same as the Word CRC template)
% - added example for generating copyright line in abstract

%-----------------------------------------------------------------------------------

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                          %%
%% Important note on usage                                  %%
%% -----------------------                                  %%
%% This file should normally be compiled with PDFLaTeX      %%
%% Using standard LaTeX should work but may produce clashes %%
%%                                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
%% The 'procedia' option causes ecrc to approximate to the Word template
\documentclass[3p,times,procedia]{elsarticle}
\flushbottom

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
\usepackage[bookmarks=false]{hyperref}
    \hypersetup{colorlinks,
      linkcolor=blue,
      citecolor=blue,
      urlcolor=blue}
%\usepackage{amsmath}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Procedia Computer Science}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{Author name}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
%\jnltitlelogo{Computer Science}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols

\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{authoryear}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}
%\usepackage{harvard}
% put your own definitions here:x
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}
\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{The 9th International Workshop on Agent-based Mobility, Traffic and Transportation Models, \\ (ABMTRANS) March 23 - 26, 2021, Warsaw, Poland}%
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}
%% \dochead can also be used to include a conference title, if directed by the editors
%% e.g. \dochead{17th International Conference on Dynamical Processes in Excited States of Solids}

\title{Insert here, the title of your paper, Capitalize first letter}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[a]{First Author} 
\author[b]{Second Author}
\author[a,b]{Third Author\corref{cor1}}

\address[a]{First affiliation, Address, City and Postcode, Country}
\address[b]{Second affiliation, Address, City and Postcode, Country}

\begin{abstract}
%% Text of abstract
Insert here your abstract text.
\end{abstract}

\begin{keyword}
Type your keywords here, separated by semicolons ;

%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}
\cortext[cor1]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000.}
\end{frontmatter}

%\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000.}
\email{author@institute.xxx}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text

%\enlargethispage{-7mm}
\section{Main Text}
\label{main}

Here introduce the paper, and put a nomenclature if necessary, in a box with the same font size as the rest of the paper. The paragraphs continue from here and are only separated by headings, subheadings, images and formulae. The section headings are arranged by numbers, bold and 10 pt. Here follows further instructions for authors.

\begin{nomenclature}
\begin{deflist}[A]
\defitem{A}\defterm{radius of}
\defitem{B}\defterm{position of}
\defitem{C}\defterm{further nomenclature continues down the page inside the text box}
\end{deflist}
\end{nomenclature}

\subsection{ Structure}
Files must be in LaTeX format only and should be formatted for direct printing, using the CRC LaTeX template provided. Figures and tables should be embedded and not supplied separately.

Please make sure that you use as much as possible normal fonts in your documents. Special fonts, such as fonts used in the Far East (Japanese, Chinese, Korean, etc.) may cause problems during processing. To avoid unnecessary errors you are strongly advised to use the `spellchecker' function of TeX Editor. Follow this order when typing manuscripts: Title, Authors, Affiliations, Abstract, Keywords, Main text (including figures and tables), Acknowledgements, References, Appendix. Collate acknowledgements in a separate section at the end of the article and do not include them on the title page, as a footnote to the title or otherwise.

Bulleted lists may be included and should look like this:
\begin{itemize}[]
\item First point
\item Second point
\item And so on
\end{itemize}

Ensure that you return to the `Els-body-text' style, the style that you will mainly be using for large blocks of text, when you have completed your bulleted list.

Please do not alter the formatting and style layouts which have been set up in this template document. As indicated in the template, papers should be prepared in single column format suitable for direct printing onto paper with trim size $192 \times 262$ mm. Do not number pages on the front, as page numbers will be added separately for the preprints and the Proceedings. Leave a line clear between paragraphs. All the required style templates are provided in the file ``LaTeX Template'' with the appropriate name supplied, e.g. choose 1. Els1st-order-head for your first order heading text, els-abstract-text for the abstract text etc.

\subsection{ Tables}

All tables should be numbered with Arabic numerals. Every table should have a caption. Headings should be placed above tables, left justified. Only horizontal lines should be used within a table, to distinguish the column headings from the body of the table, and immediately above and below the table. Tables must be embedded into the text and not supplied separately. Below is an example which the authors may find useful.

\begin{table}[h]
\caption{An example of a table.}
\begin{tabular*}{\hsize}{@{\extracolsep{\fill}}lll@{}}
\toprule
An example of a column heading & Column A ({\it{t}}) & Column B ({\it{t}})\\
\colrule
And an entry &   1 &  2\\
And another entry  & 3 &  4\\
And another entry &  5 &  6\\
\botrule
\end{tabular*}
\end{table}

%\enlargethispage{12pt}

\subsection{ Construction of references}

References must be listed at the end of the paper. Do not begin them on a new page unless this is absolutely necessary. Authors should ensure that every reference in the text appears in the list of references and vice versa. Indicate references by \cite{Massimo2011} or \cite{Massimo2012} or \cite{Thomas2015} in the text.

Some examples of how your references should be listed are given at the end of this template in the `References' section, which will allow you to assemble your reference list according to the correct format and font size.

Reference generation by using bibliography style commands for LaTeX template only.

The author may use ``elsarticle-harv.bst'' as per the style required in document. The sample bib file could be referred. 
If the author may using bibstyle for providing references author must comment the bibliography section in TeX file, Bibtex will generate the reference automatically.

If the author may not able to view the references in output same could be done by copying the bibliography section from ``filename.bbl'' file and paste in TeX file.

\subsection{Section headings}
Section headings should be left justified, bold, with the first letter capitalized and numbered consecutively, starting with the Introduction. Sub-section headings should be in capital and lower-case italic letters, numbered 1.1, 1.2, etc,~and left justified, with second~and subsequent lines indented. All headings should have a minimum of two text lines after them before a page or column break.
Ensure the text area is not blank except for the last page.

\subsection{General guidelines for the preparation of your text}
Avoid hyphenation at the end of a line. Symbols denoting vectors and matrices should be indicated in bold type. Scalar variable names should normally be expressed using italics. Weights and measures should be expressed in SI units. All non-standard abbreviations or symbols must be defined when first mentioned, or a glossary provided.

\subsection{File naming and delivery}
Please title your files in this order `procedia acronym\_conference acronym\_authorslastname'.  Submit both the source file and the PDF to the Guest Editor.

Artwork filenames should comply with the syntax ``aabbbbbb.ccc'', where:
\begin{itemize}
\item a = artwork component type
\item b = manuscript reference code
\item c = standard file extension

Component types:
\item gr = figure
\item pl = plate
\item sc = scheme
\item fx = fixed graphic
\end{itemize}

\subsection{Footnotes}
Footnotes should be avoided if possible. Necessary footnotes should be denoted in the text by consecutive superscript letters\footnote{Footnote text.}. The footnotes should be typed single spaced, and in smaller type size (8 pt), at the foot of the page in which they are mentioned, and separated from the main text by a one line space extending at the foot of the column. The `Els-footnote' style is available in the ``TeX Template'' for the text of the footnote.

Please do not change the margins of the template as this can result in the footnote falling outside printing range.

\section{Illustrations}
All figures should be numbered with Arabic numerals (1,2,3,$\,\ldots.$). Every figure should have a caption. All\break photographs, schemas, graphs and diagrams are to be referred to as figures. Line drawings should be good\break quality scans or true electronic output. Low-quality scans are not acceptable. Figures must be embedded into the text and not supplied separately. In MS word input the figures must be properly coded. Preferred format of figures are PNG, JPEG, GIF etc. Lettering and symbols should be clearly defined either in the caption or in a legend provided as part of the figure. Figures should be placed at the top or bottom of a page wherever possible, as close as possible to the first reference to them in the paper. Please ensure that all the figures are of 300 DPI resolutions as this will facilitate good output.

The figure number and caption should be typed below the illustration in 8 pt and left justified [{\bfseries\itshape Note:} one-line captions of length less than column width (or full typesetting width or oblong) centered]. For more guidelines and\break information to help you submit high quality artwork please visit: http://www.elsevier.com/artworkinstructions\break Artwork has no text along the side of it in the main body of the text. However, if two images fit next to each other, these may be placed next to each other to save space. For example, see Fig.~1.
\begin{figure}[t]\vspace*{4pt}
%\centerline{\includegraphics{fx1}\hspace*{5mm}\includegraphics{fx1}}
\centerline{\includegraphics{gr1}}
\caption{(a) first picture; (b) second picture.}
\end{figure}

\section{Equations}
Equations and formulae should be typed in MathType, and numbered consecutively with Arabic numerals in parentheses on the right hand side of the page (if referred to explicitly in the text). They should also be separated from the surrounding text by one space
\begin{equation}
\begin{array}{lcl}
\displaystyle X_r &=& \displaystyle\dot{Q}^{''}_{rad}\left/\left(\dot{Q}^{''}_{rad} + \dot{Q}^{''}_{conv}\right)\right.\\[6pt]
\displaystyle \rho &=& \displaystyle\frac{\vec{E}}{J_c(T={\rm const.})\cdot\left(P\cdot\left(\displaystyle\frac{\vec{E}}{E_c}\right)^m+(1-P)\right)}
\end{array}
\end{equation}

\section{Online license transfer}
All authors are required to complete the Procedia exclusive license transfer agreement before the article can be published, which they can do online. This transfer agreement enables Elsevier to protect the copyrighted material for the authors, but does not relinquish the authors' proprietary rights. The copyright transfer covers the exclusive rights to reproduce and distribute the article, including reprints, photographic reproductions, microfilm or any other reproductions of similar nature and translations. Authors are responsible for obtaining from the copyright holder, the permission to reproduce any figures for which copyright exists.

\section*{Acknowledgements}

Acknowledgements and Reference heading should be left justified, bold, with the first letter capitalized but have no numbers. Text below continues as normal.

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

\appendix
\section{An example appendix}
Authors including an appendix section should do so before References section. Multiple appendices should all have headings in the style used above. They will automatically be ordered A, B, C etc.

\subsection{Example of a sub-heading within an appendix}
There is also the option to include a subheading within the Appendix if you wish.
%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%The citation must be used in following style: \cite{article-minimal} \cite{article-full} \cite{article-crossref} \cite{whole-journal}.
%% References with BibTeX database:

%\bibliography{xampl}
%\bibliographystyle{elsarticle-harv}

%% Authors are advised to use a BibTeX database file for their reference list.
%% The provided style file elsarticle-num.bst formats references in the required Procedia style

%% For references without a BibTeX database:

 \begin{thebibliography}{}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

\bibitem{Massimo2011}
{F}ilippini, Massimo, and Lester C. Hunt. (2011) ``Energy demand and
energy efficiency in the OECD countries: a stochastic demand frontier
approach." {\it Energy Journal} {\bf 32} (2): 59--80.
\bibitem{Massimo2012}
Filippini, Massimo, and Lester C. Hunt. (2012) ``US residential
energy demand and energy efficiency: A stochastic demand frontier
approach." {\it Energy Economics} {\bf 34} (5): 1484--1491.
\bibitem{Thomas2015} 
Weyman-Jones, Thomas, J\'{u}lia Mendon\c{c}a Boucinha, and Catarina
Feteira In\'{a}cio. (2015) ``Measuring electric energy efficiency in
Portuguese households: a tool for energy policy." {\it Management of Environmental Quality: An International Journal} {\bf 26} (3): 407--422.
\bibitem{} 
Saunders, Harry (2009) ``Theoretical Foundations of the Rebound Effect'', in Joanne Evans and Lester Hunt (eds) {\it International Handbook on the Economics of Energy}, Cheltenham, Edward Elgar
\bibitem{} 
Sorrell, Steve (2009) ``The Rebound Effect: definition and estimation'', in Joanne Evans and Lester Hunt (eds) {\it International Handbook on the Economics of Energy}, Cheltenham, Edward Elgar 
 \end{thebibliography}

\clearpage

%%%% This page is for instructions only, once the article is finalize please omit the below text before creating the final PDF
\normalMode

\section*{Instructions to Authors for LaTeX template:}

\section{ZIP mode for LaTeX template:}

The zip package is created as per the guide lines present on the URL http://www.elsevier.com/author-schemas/ preparing-crc-journal-articles-with-latex for creating the LaTeX zip file of Procedia LaTeX template.  The zip generally contains the following files:
\begin{Itemize}[]\leftskip-17.7pt\labelsep3.3pt
\item ecrc.sty
\item  elsarticle.cls
\item elsdoc.pdf
\item .bst file
\item Manuscript templates for use with these bibliographic styles
\item  Generic and journal specific logos, etc.
\end{Itemize}

The LaTeX package is the main LaTeX template. All LaTeX support files are required for LaTeX pdf generation from the LaTeX template package.

{\bf Reference style .bst file used for collaboration support:} In the LaTeX template packages of all Procedia titles a new ``.bst'' file is used which supports collaborations downloaded from the path http://www.elsevier.com/author-schemas/the-elsarticle-latex-document-class

\section{Reference style used in Computer Science:}
\let\footnotesize\normalsize
\hspace*{-10pt}\begin{tabular*}{\hsize}{@{}ll@{}}
{\bf Title}&{\bf Reference style} \\[6pt]
PROCS  & 3 Vancouver Numbered
\end{tabular*}

\end{document}

%%
%% End of file `procs-template.tex'.

### 2020-12-30 src/paper.tex
checkpoint: bibs working, abstract and intro written

@@
-% sage_latex_guidelines.tex V1.20, 14 January 2017
+\documentclass[3p,times,procedia]{elsarticle}
+\flushbottom
 
-\documentclass[Afour,sagev,times]{sagej}
+%% The `ecrc' package must be called to make the CRC functionality available
+\usepackage{ecrc}
+\usepackage[bookmarks=false]{hyperref}
+    \hypersetup{colorlinks,
+      linkcolor=blue,
+      citecolor=blue,
+      urlcolor=blue}
+%\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage[utf8]{inputenc}
 
-\usepackage{moreverb,url}
+%% set the volume if you know. Otherwise `00'
+\volume{00}
 
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
+%% set the starting page if not 1
+\firstpage{1}
 
-% -- to allow highlight changed sections in yellow:
-\usepackage{color,soul}
+%% Give the name of the journal
+\journalname{Procedia Computer Science}
 
-\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
-T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+%% Give the author list to appear in the running head
+%% Example \runauth{C.V. Radhakrishnan et al.}
+\runauth{W. Charlton et al.}
 
-\def\volumeyear{2020}
+%% Give the abbreviation of the Journal.
+\jid{procs}
 
-%%Vancouver (numbered)
-\bibliographystyle{SageV}
-%\bibliography{paper.bib}
+\usepackage{amssymb}
+
+% if you have landscape tables
+\usepackage[figuresright]{rotating}
+
+% add words to TeX's hyphenation exception list
+%\hyphenation{author another created financial paper re-commend-ed Post-Script}
+
+% declarations for front matter
 
 \begin{document}
+\begin{frontmatter}
 
-\runninghead{Charlton and Laudan}
+\dochead{The 9th International Workshop on Agent-based Mobility, Traffic and Transportation Models, \\ (ABMTRANS) March 23 - 26, 2021, Warsaw, Poland}%
 
-\title{A Web-Based Data Visualization Platform for MATSim}
+\title{Open-Source Web-Based Visualizer for \\ Dynamic-Response Shared Taxi Simulations}
 
-\author{Billy Charlton\affilnum{1} and Janek Laudan\affilnum{1}}
+\author[a]{William Charlton}
+\author[a]{Gregor Leich}
+\author[a]{Ihab Kaddoura}
+\author[a]{Kai Nagel}
 
-\affiliation{\affilnum{1}Technische Universit\"at Berlin}
+\address[a]{Technische Universität Berlin, Chair of Transport Systems Planning and Transport Telematics, Straße des 17. Juni 135, 10623 Berlin, Germany}
 
-\corrauth{Billy Charlton, Technische Universit\"at Berlin,
-Transport Systems Planning and Transport Telematics,
-Secretariat SG 12,
-Salzufer 17-19, 10587 Berlin, Germany}
+%% ----- ABSTRACT -----
+\begin{abstract}
 
-\email{charlton@tu-berlin.de}
+We describe a unique, web-based data visualization portal developed for use by researchers and public transit agencies investigating future shared-taxi fleet scenarios.
+Augmenting or even replacing fixed-route transit lines with automated, connected, shared taxi fleets may be a desirable alternative in less-densely developed areas.
+The MATSim agent-based transport microsimulation model is used to study scenarios including status quo; dynamically-dispatched fleets with drivers; and fully autonomous fleets.
+This paper focuses on a data visualization portal which includes many interactive views, such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more.
+The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g.
+school pickup/drop-offs or commute trips connecting to rail stations.
+The tool is in operation for several small cities and rural regions in Germany and was successfully used as an outreach tool in public meetings.
+In addition, developers of the MATSim DRT extension found the visualizations particularly useful for debugging both the algorithms and the scenario definitions.
+All code is open source and, while this specific study has a very specific use case, the portal has an extensible design that could be modified for other purposes.
 
-\begin{abstract}
-There are many tools available for analyzing MATSim transport
-simulation results, both open-source and commercial. This research
-builds a new open-source visualization platform for MATSim outputs that
-is entirely web-based. After initial experiments with many different web
-technologies, a client/server platform design emerges which leverages
-the advanced user interface capabilities of modern browsers on the
-front-end, and relies on back-end server processing for more processor-intensive
-tasks. The initial platform is now operational and includes several
-aggregate-level visualizations including origin/destination flows,
-transit supply, and emissions levels; as well as a fully disaggregate
-traffic animation visualization. These visualizations are general
-enough to be useful for various projects. Further work is needed
-to make them more compelling and the platform more useful for practitioners.
 \end{abstract}
 
-\keywords{MATSim, data visualization, agent-based simulation}
+\begin{keyword}
+shared taxis; data visualization; matsim; public outreach
+\end{keyword}
 
-\maketitle
+%\correspondingauthor[*]{Corresponding author. Tel.: +0-000-000-0000 ; fax: +0-000-000-0000.}
+\cortext[cor1]{Corresponding author. Tel.: +49 30 314-23308 ; fax: +49 30 314-26269.}
+\end{frontmatter}
 
+\email{charlton@vsp.tu-berlin.de}
+
+%% ----- INTRODUCTION -----
 \section{Introduction}
+\label{introduction}
 
-MATSim\cite{R1} is an open-source framework for implementing large-scale agent-based transport simulations.
-MATSim is widely used for transportation research in academic settings, and is gaining momentum as a tool ready for practice in real-world planning contexts.
+Augmenting or replacing fixed-route transit lines with automated, connected, shared taxi fleets may be a desirable alternative for transit agencies in less-densely developed areas.
+The economics of fixed-route transit are difficult to square with sparse development patterns [*], and the advent of private ride-hailing services adds the "shared taxi" concept to people's travel options.
+The same technologies used by private ride-hailing, however, can also be modified and applied to future transit services.
 
-There are many tools available for analyzing MATSim results, both
-open-source and commercial. Typically, analysts can choose either
-the free tool OTFVis\cite{R2} or the commercial software
-Via \cite{R3}, both of which are desktop software packages requiring installation as well as a fair amount of technical acumen to operate.
-Alternatives to these tools include the more general-purpose desktop mapping software packages such as QGis and ArcGIS, or statistical software packages, again all of which require installation and expertise to use.
+In this study, a new dynamic-response shared taxi service, or "DRT", is examined as an alternative approach to serving the population of rural areas and less-dense portions of urban regions.
+Dynamic-response transit is analyzed using the MATSim agent-based transport microsimulation model.
+MATSim is an extensible, activity- and multi-agent based transport simulation, which enables the simulation of large scale scenarios.
+\cite{MATSimBook}.
+Crucially, MATSim simulates passenger transport in both private and public modes \cite{ZiemkeEtAl2019OpenBerlinScenario}.
 
-As MATSim moves from the confines of academic research to a more
-public-facing role, a notable gap is apparent: there are no web-based interactive tools available for disseminating MATSim data and results.
-This creates a challenge for using MATSim in public policy settings: the only people who can meaningfully examine and explore results are those who have extensive technical knowledge and access to the specialized software and large datasets involved.
+The existing methodology behind using MATSim to simulate DRT is well-described by [*] and is not the focus of this paper.
+Rather, in the course of the effort to develop, analyze and compare various DRT scenarios, a need arose to create some truly unique data visualizations that could assist in improving and displaying the DRT scenarios themselves.
+A web-based portal is the result of this research.
+The portal includes many interactive views, such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more.
+The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g.
+school pickup/drop-offs or commute trips connecting to rail stations.
 
-This research explores one way to fill this gap: building an open
-web-based data visualization platform which is specifically designed
-to complement MATSim.
+The tool is now in operation for several small cities and rural regions in Germany, and was successfully used as an outreach tool in online public meetings.
+Project partners at transit agencies are interested in understanding the detailed implications of the DRT simulations, and without a visual, interactive component, that would be extremely difficult to provide.
 
-\section{Motivation}
+Furthermore, developers of the MATSim DRT extension found the visualizations particularly useful for debugging both the algorithms and the scenario definitions.
+The latest implementions of MATSim include extensive DRT capabilities whose methods are continually being optimized and improved.
 
-The rapid advancement of Internet browsing technologies in the last
-five years has enabled the web browser to do things much more
-"application"-like than ever before: background processing,
-three-dimensional rendering using \hl{graphics card (GPU)} acceleration, offline
-support, and more. The combination of these technologies and their
-standard implementations on every popular hardware type and operating system now makes the web a very compelling platform.
+All code for MATSim and the visualization portal is open source and, while this specific study has a quite esoteric use case, the portal has an extensible design that could be modified for other purposes.
 
-For MATSim research, the question is: could a web browser really
-be useful for exploring and delivering results when the datasets
-are so large? Answering this question is the primary motivation for
-this research. Essentially, has the web become powerful enough for MATSim?
+%% ----- EXISTING TOOLS AND RESEARCH -----
+\section{Motivation: The State of Existing Tools}
+\label{motivation}
 
-Currently, analysis of MATSim outputs ends up in research reports, \hl{Adobe (PDF)} documents, video screen-recordings, and presentations.
-An online dashboard of results that a user could explore and manipulate would not only be more interactive, but might also reveal findings that the original analysts hadn't anticipated.
+- Existing tools are incapable of visualizing the outputs of the DRT module of MATSim in an efficient manner
+- Authors of the DRT modules of MATSim expressed an unfulfilled need to debug the code
+- Previous research on web-based MATSim visualization provides a proof-of-concept and extensible open-source platform on which to build DRT-specific visualizations; an obvious extension!
+- Agent-based microsimulation is quite complex and the reams of data and extensive details are not appropriate for public outreach and stakeholder meetings.
+PDF reports and powerpoint slides are the norm, but there is an opportunity here to build something truly interactive, possibly an order of magnitude more useful.
 
+%% ----- REQUIREMENTS -----
 \section{Requirements}
-The research team at Technische Universit\"at Berlin had several
-"blank slate" discussions before any code was written: meaning,
-if we could start at the very beginning and design something completely web-based and open, what would the bare minimum requirements be for it to be truly useful?
-The following requirements emerged from those discussions.
+\label{requirements}
 
-\subsection{Requirement 1: Web browser-based}
+- Open source code so others may build upon it and extend it
+- Works in any modern browser. IE11 and small players, no. (pull from other doc)
+- Must somehow manage massive output datasets from MATSim runs. Browsers can't load 800MB files support massive datasets; requires some sort of output post-processing and storage of results
+- Visually compelling and easy to organize results for various scenarios
+- Public facing scenarios and private debugging areas
 
-Given the above-stated motivation and hypothesis that the modern web
-platform is ready for large-scale visualization tasks, the most obvious requirement is that the product of this research must work with any modern web browser.
+%% ----- EXPERIMENTS -----
+\section{The Visualization Platform}
+\label{platform}
 
-Several specific web technologies developed and made widely available in recent years enable us to perform this research: HTML 5, CSS 3, WebGL, ECMA Script 6, and Web Workers.
-Briefly, these technologies are:
+Platform
 
-\begin{enumerate}
+- MatHub platform had extensive back-end server infrastructure (six server images for user authentication, file storage, post-processing, real-time animations). But it sucked
+- MatHub front-end used MapBox API, but size of datasets made all but the simplest of scenarios difficult or impossible to visualize within browser memory constraints.
 
-\item[(i)] \textbf{HTML 5} improves and standardizes the "document model" of what constitutes a web page and how it is specified.
-\item[(i)] \textbf{CSS 3} is a styling language that enables fine-grained layout and styling of individual elements on a page.
-CSS 3 defines in a consistent, standard way the details of elements including color, size, layout, and animation of page elements.
-\item[(ii)] \textbf{WebGL} provides in-browser support for the 3D-accelerated graphics capabilities of modern machines.
-\item[(iv)] \textbf{ECMAScript 6} is an updated (2015) version of the venerable Javascript scripting language that has been part of the web platform since the early 1990's.
-Recent versions of Javascript eliminate the more problematic aspects of the language and make it easier for developers to create bug-free, efficient code.
-\item[(v)] \textbf{Web Workers} are a recent (2013) addition to the web platform that allow background processing for long-running tasks.
-Before Web Workers, there was no way to run truly multi-threaded code inside a browser.
+- Covid-19 portal threw out everything except centralized file storage, and relied on the analyst to run standardized post-processing scripts and store those smaller datasets.
+First experiments with ThreeJS were successful but very time consuming to write
+- deck.gl proved far more performant than MapBox, also open source, and easier to write custom "shaders" using deck.gl examples
 
-\end{enumerate}
+DRT animation
 
-A complication in web development is that the major web browser vendors implement these technologies on their own timelines, some much more rapidly than others.
-Further complicating things is the reality that end users do not always upgrade their browsers frequently (or at all).
-This creates a landscape where there is a technology adoption curve with a very long "tail".
-Thus, developers of every web site need to make a conscious decision about where to
-draw the line -- choosing necessary technologies for their site to operate correctly, while knowing that some users with older browsers will either have a sub-optimal experience or no access to the site at all.
+- Initially just the vehicle animations
+- Analysts requested color by occupancy
+- origin/destination "flyovers"
 
-For this research, we are deliberately exploring the \textit{latest standard} web technologies, with the expectation that access to these technologies will become more and more common in the future.
-Thus, we are targeting the most recent versions of modern web browsers as of 2019, including Google Chrome, Mozilla Firefox,
-and Apple Safari.
-All three browsers fully support the above-listed technologies, and importantly, all three auto-update automatically, ensuring that most users of those browsers stay current as these technologies evolve.
-\hl{This does leave out older browsers such as Internet Explorer and Edge, as well as browsers with small market share.
-Supporting those browsers would require much more deliberate and time-consuming testing, which while not impossible, is currently out of scope.
-For some government users this is acknowledged as possibly problematic, since use of Internet Explorer 11 is still mandated at some agencies.}
+%% ----- DRT VISUALIZATIONS -----
+\section{Dynamic-Response Taxi Visualizations}
+\label{drtviz}
 
-\subsection{Requirement 2: Open source}
+{Strippgen2016}
+Back-end consists only of file storage.
+Our research team has an existing "Subversion" server which handles large files efficiently and retains file version history.
+Other file storage solutions such as cloud-based AWS etc could also be implemented (but weren't) -- all file management goes through a single File API which could have multiple back-ends.
 
-The entire project, including all front-end (browser) and back-end (server) code, must be fully open-source.
+Front end code is a "single page app" that can be hosted on any web server platform. We are using Github Pages because it is free and fast.
 
-No proprietary or closed licensing schemes were considered, primarily because excellent proprietary visualization packages for MATSim already exist.
-Creating a competing product would be duplicative and unnecessary, and would not further the research goal of determining whether web-based technology is now advanced enough to work with MATSim outputs.
-The goal of this research is not to replace existing proprietary solutions, but rather to complement them.
+Web portal front-end has a simple text-based configuration file which lists the scenarios to be publicly viewable
+- Analyst runs agent-based model and post-processing scripts to create road/transit network and DRT animation files
+- etc
 
-The software developed as part of this research is licensed entirely with the GNU General Public License version 3, commonly known as the \textit{GPL v3}\cite{R4}.
+%% ----- RESULTS -----
+\section{Results}
+\label{results}
 
-This matches the license of MATSim itself.
-Several other open-source licenses were considered, including the MIT License and the Apache Public License, but the benefit of sharing a common license with MATSim outweighs any perceived benefits of switching to other open licenses \hl{The main benefit of those other open licenses is compatibility with commercial software.
-Given its narrow scope, this research is not likely to be of interest to existing commercial vendors.
-Furthermore, the research team does not intend to turn this into a commercial product.
-Thus the GPL license is a good fit.}
+- It works!
+- Website
+- Six public scenarios
+- Gregor's findings
+- Transit agency public outreach meetings
 
-\subsection{Requirement 3: Use good defaults, with minimal configuration, and be opinionated}
+FROM GREGOR
 
-Since its inception, the web platform has had a relentless focus on simplicity and smooth user onboarding.
-Users are accustomed to being immediately familiar with a site -- often within seconds of their first interaction.
-Because of this expectation, it is critical that this research follow current best practices
-for user interface (UI) and user experience (UX).
-Specifically, that means using familiar UI paradigms such as navigation bars and breadcrumbs, separating configuration from usage, limiting settings and options to the bare minimum, and being "opinionated", i.e.
-encouraging a correct way to accomplish a task.
+Not sure whether this helps, but is what comes to my mind: The drt viz helped us understand better how the drt code works and find some issues, e.g.
+when the system is congested weird things can happen and a large group of vehicles with one passenger per vehicle only moves from the same start to the same destination at the same time (in an attempt to save each passenger a few seconds it serves them separately and wastes vehicles which then are missing for the next requests).
+It is really nice to have requests and vehicle occupancy shown, Via did only the latter after some pre-processing.
+Other than that the website obviously made the avoev workshops easier.
+Instead of only sharing our screen we could hand out a website where people could click themselves.
+Unfortunately it seemed that those listeners did not use the website a lot and rather kept listening to us.
+Some hypotheses: Maybe because sometimes a group of people shared one computer and their internet connection was bad, maybe because they are not used to that kind of interaction (maybe they always just listen to others' presentations unless they present themselve and discuss verbally later).
+For us the other plots like pax/link or veh/link are not entirely new, some used QGis to produce those.
+However maybe the website forced us to actually produce all those plots systematically instead of an "we could do that plot, but maybe it's not necessary, let's look into something else or raw data only to avoid the hassle of setting up QGis..."
 
-This approach is dissimilar to some data exploration tools (e.g., QGis and Via) where extreme configurability is emphasized.
-Rather than providing myriad options for details such as scales and color ramps, our research focuses on choosing good defaults and determining whether that is sufficient for the software to be
-useful.
+%% ----- CONCLUSIONS -----
+\section{Conclusions and Outlook}
+\label{conclusions}
 
-\subsection{Requirement 4: An extensible platform}
+This is the platform we've been trying to build for three years. It's efficient, extensible, and open source.
 
-Every data visualization use case is different; there is no way to anticipate how the tool will be used.
-If the platform is too generic, it will be not at all useful.
-Conversely, if only hard-coded visualizations are created for specific projects, it will be relegated to "demo-ware", meaning it is a successful technology demonstration but not actually useful for real users.
+Further enhancements to the underlying platform and to the DRT visualization capabilities are expected.
 
-To fulfill this requirement the software platform will need to be extensible: basic capabilities and templates will be provided, but a user with some level of coding skill should be able to create new visualizations that are not anticipated by the researchers.
+Plugin architecture means that other javascript developers can write plugins for their own agent-based models and their own use cases.
 
-\section{Initial experiments}
+\section{Acknowledgements}
+This research was funded in part by the German Federal Ministry of Transport and Digital Infrastructure (funding number 16AVF2160)
 
-It is no exaggeration to state that the Javascript code library ecosystem is extremely, enormously large.
-Thousands of libraries and packages are available on a common centralized Javascript repository known as "NPM", and there are often multiple packages that do similar things.
-As a developer, one must assess and select from these packages or choose to solve a problem by writing code by hand.
-Of course these libraries are of varying levels of popularity and quality.
+\bibliography{vsp,ref,book}
+\bibliographystyle{elsarticle-harv}
 
-Based on the requirements laid out above, some initial experiments were carried out to assess various approaches before committing to a technology stack.
+% \begin{thebibliography}{1}
 
-\subsection{Visualizing time-dependent data on a geographic map}
+%   \bibitem{R1}
+%   Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016). \textit{The multi-agent transport simulation MATSim} (p. 618). London: Ubiquity Press.
 
-Two popular web-based Javascript libraries were tested for displaying geographic data; Leaflet and Mapbox GL.
-A simple test case comprised of MATSim simulation outputs was developed, with the goal of displaying aggregated roadway link volumes by time of day.
-
-Leaflet (\url{leafletjs.com}) is very popular and its application programming interface (API) is a bit simpler than that of Mapbox GL.
-Leaflet uses background map "tiles" at specific zoom levels, and layers data on top of those base maps.
-With small networks (we tested Cottbus, Germany, a small city of 100,000 inhabitants) Leaflet performed well, but medium-sized and large-sized networks with many elements visible at once suffered from noticeable performance degradation.
-This was problematic, as this was the simplest use-case envisioned.
-
-Mapbox GL (\url{mapbox.com}) fared much better, apparently better-suited to displaying large datasets with many visible features simultaneously.
-In addition, Mapbox GL's use of 3D vector graphic mapping instead of preset tiles made for a much more pleasing user experience, with smooth animations between zoom levels and better background processing during page loads.
-For these reasons, Mapbox GL was chosen as the base map for the remaining geographic visualizations.
-
-\subsection{Visualizing non-geographic data}
-
-There are hundreds of data visualization libraries available for the web which provide ways to produce charts and plots of varying complexity.
-Our requirement of using open-source code narrows the field considerably.
-
-After experimenting with several packages including D3, Raphael, Morris and others, the package Vega-Lite\cite{R5} exhibited many of the characteristics desired.
-Notably, Vega-Lite follows a declarative syntax known as the "grammar of graphics", as popularized by Wilkinson\cite{R6}.
-This grammar allows concise description of the meaningful components of a graphic.
-An added benefit is that graphs and charts can be downloaded in image format or in Vega source format, which is helpful for other researchers wanting to learn how to use the data format.
-
-\subsection{Dealing with large datasets}
-
-MATSim network files are small enough to fit in \hl{computer memory (RAM)}, but MATSim plan files and event files can be much larger than RAM, necessitating careful consideration about how to handle them.
-
-Modern browsers allow access via API to a data storage area that is unique per hostname, i.e.
-http://mysite.com is allowed some storage on the local machine.
-Initial experiments revealed that each browser vendor implements this storage differently, with very strict limits on the absolute amount storage available, sometimes based on how much free space remains on the user's machine.
-It became apparent that this local browser-based storage would not be sufficient for MATSim outputs.
-Running a local file-server process would allow browser access to a specific folder on the machine, which might be nice for advanced analysts but violates the research goal of being fully web-based on the client end.
-Thus, a client-server paradigm emerges as the only truly viable alternative, and indeed this is how most websites operate today: the browser is the front-end to the heavier processing and storage tasks that happen on someone else's server.
-Note that "someone else's server" is usually referred to as "the cloud".
-
-\section{Platform architecture}
-
-A client/server architecture was chosen for this research based on the initial experiments described above.
-
-The research team authored back-end server software for file storage, user authentication, and data pre-processing.
-Due to space considerations, this paper does not delve into the details of those components.
-Suffice it to mention that the front-end communicates with them to establish what resources a user has access to, and provides an application programming interface, or API, with which to query and fetch available files and resources.
-The code for those servers is also open source, and may be the subject of future papers.
-
-The front-end architecture has several interacting components, described below.
-
-\subsection{"Vue" Single Page Application}
-
-The primary framework used to build the application is known as "VueJS" or Vue (available at \url{vuejs.org}).
-Vue is a framework for building interactive user interfaces on the web; it provides the glue between the items a user clicks and the code that runs when they do so.
-Vue provides many services which allow a web page to behave more like a full-featured application, including state management, routing between different URLs, and componentization of code in a way that encourages code reuse and loose coupling.
-Vue depends on Javascript, which means it \hl{does not work for users} who have disabled scripting in their browser.
-
-Vue is most often used to write so-called "single page applications" which are websites that behave like desktop applications.
-Most large, popular websites such as Github, Twitter, and Facebook all employ this paradigm, meaning the site handles page transitions and URLs as if they are all in one common namespace, and the user thinks less about visiting URLs and more about navigation through visual components to accomplish tasks.
-This matches our use case.
-
-Vue components each encapsulate the three elements required for the modern web: HTML layout, Javascript code, and CSS formatting.
-Components only interact with each other through well-defined pathways of properties and events, which greatly improves debuggability.
-
-\subsection{Build system}
-
-The build system of a modern web application is fairly complex and the Javascript ecosystem changes rapidly.
-After numerous iterations, the current build system comprises a series of individual tools that all work together to produce the final assets that get sent to a user's browser.
-Those tools include the Vue command line interface (CLI), the NPM package manager, webpack, babel, and TypeScript.
-
-Notably, the initial codebase was migrated to the Typescript language midway through development, as the benefits of a strongly-typed language were perceived to be worth the development time.
-TypeScript is a separate language from JavaScript, and enforces type annotations for variables and adds additional features such as enumerations.
-The TypeScript compiler then converts TypeScript code into ECMAScript 6 JavaScript, which can be run in the browser (as JavaScript is the only scripting language that browsers support).
-
-\subsection{Visualization plug-ins}
-
-One of the main requirements of this research is to produce a system where new visualizations can be produced rapidly and added into the existing framework to generate new capabilities.
-
-The Vue component architecture enables this.
-To create a new visualization, a developer copies an existing "blank" visualization template and gives it a new name, specifies the file inputs and parameters required, and then uses the above-described libraries to modify the code per their needs.
-
-This currently requires ample coding skill in Javascript; it is not a system that is point-and-click like an online data exploration tool.
-Experience with other similarly-designed platforms suggests that software-minded analysts or modelers would be able to extend the platform, but typical end users will not.
-
-\section{Results: the current state of the tool}
-
-A working instance of the platform is now online and available at \url{viz.vsp.tu-berlin.de}.
-Sample datasets are uploaded, and pre-built visualizations are publicly accessible, as a demonstration of the platform's current state.
-There is also a user login system so that internal researchers can extend and experiment with the system, without exposing data or work-in-progress to the public.
-
-Basic user, project, and file management capabilities are operational. This includes grouping files by model run or by other user-defined tags, as can be seen in Figure~\ref{F1}.
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-1.png}
-\caption{Project dashboard}\label{F1}
-\end{figure}
-
-See the following screenshots in Figures~\ref{Fod}-~\ref{Fanim} for examples of the current state of the user interface.
-Note of course that the tool demos better live than in screenshots, as the whole point of this research is to develop an interactive tool.
-
-Several types of aggregate visualizations are developed:
-
-\begin{enumerate}
-
-\item[(i)] Origin/destination flows between aggregate areas, so-called "spider diagrams" (Figure~\ref{Fod})
-\item[(i)] Link flows, by time of day and mode (Figure~\ref{Flinkvols})
-\item[(iii)] Transit supply explorer, which displays all transit routes and allows the user to see which routes serve specific stops and links. (Figure~\ref{Ftransit})
-\item[(iv)] Sankey diagrams, which can be used to depict changes/flows between between scenarios across multiple choices, such as shifts in mode between two scenarios (Figure~\ref{Fsankey})
-\item[(v)] Emissions levels on a geographic hexagonal grid basis (Figure~\ref{Femissions})
-
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-2.png}
-\caption{Origin/Destination flows}\label{Fod}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-3.png}
-\caption{Link volume summary}\label{Flinkvols}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-4.png}
-\caption{Transit depiction}\label{Ftransit}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-5.png}
-\caption{Mode shift diagram}\label{Fsankey}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-6.png}
-\caption{Emissions plot}\label{Femissions}
-\end{figure}
-
-In addition, one disaggregate animation is available:
-
-\begin{enumerate}
-
-\item[(i)] A vehicle flow simulation, showing individual vehicle agents in real-time on the network. (Figure~\ref{Fanim})
-
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[width=3.25in]{fig-7.png}
-\caption{Vehicle animation}\label{Fanim}
-\end{figure}
-
-\subsection{Workflow feedback}
-
-User interface design (UI) and user experience (UX) have their own entire fields of research.
-Initial prototypes of the tool simply did not meet the needs of test users.
-Problems included: difficult discerning which files were which; inability to efficiently use the model run tagging feature (in which a user could mark sets of files as belonging to a particular MATSim run); separate user logins causing data to be inaccesible to team members working on the same projects (resulting in everyone using a common "team" login, against best practices); private projects "leaking" onto the public website; and myriad usability bugs in the data visualizations themselves.
-
-These usability problems were eventually solved by surveying other technical websites which organize and present large amounts of data.
-One website in particular, \url{Github.com}, was found to be well-liked by test users and has a similar hierarchical view of data: users can belong to organizations, and both organizations and users can create projects ("repositories" on Github) which may hold large numbers of files.
-
-By adopting a project and user paradigm similar to that employed by Github, users were immediately more familiar with the site and had less to learn.
-The current site adheres to this so closely that test users often refer to the platform as "MatHub".
-
-\subsection{Performance}
-
-Even with modern hardware and the latest browsers, it is quite challenging to produce performant, visually pleasing results with disaggregate MATSim data.
-The vehicle flow simulation depends heavily on the back-end server to produce and deliver simulation "frames" to the browser in real-time, so that the browser simply has to render the data.
-
-Various levels of aggregation make MATsim data much easier to visualize, as is reflected in the number of different visualizations this research was able to produce with aggregate data in a short time frame.
-
-\textbf{Preprocessing.} The traffic animation and the emissions "hex grid" visualizations both rely on back-end server processes to preprocess the raw MATSim event outputs.
-This takes anywhere from a few seconds to many minutes, depending on the size of the simulation that was run.
-The preprocessing only needs to occur once, and thereafter the results are cached and stored on the server.
-The UI presents a helpful "still processing" message during this stage.
-Unfortunately, changing some settings such as the size of the hex-grids for the emission tool means the calculations must be redone.
-Further work is being done to make these processes run as quickly as possible.
-
-\textbf{Map layer limitations.} The Mapbox GL mapping library allows lines, points, and polygons to be displayed on top of a base map.
-There seems to be a number of map elements beyond which performance becomes very slow; various techniques can be used such as aggregating elements at further-out zoom levels to get around this.
-Another option is to use Mapbox GL only for the base map, and to use the WebGL graphics libraries to draw large numbers of elements directly on top of the map.
-This is the technique used for the traffic animation, which can easily support tens of thousands of elements (all in motion!) simultaneously.
-For additional visualizations which have large numbers of grahical elements, more research will need to be done on layering WebGL elements on top of the base map.
-
-\subsection{Mobile device support}
-
-Initially, the research targeted desktop browsers only.
-The reasoning was that analysts currently use desktop software, and it would be sufficient to complement those desktop tools with a new web-based option.
-
-However, as the tool started becoming usable by internal users, it quickly became apparent that everyone wanted some version of the visualizations to work on mobile devices, too.
-It was suggested that even a read-only visualization "viewer" for mobile would be better than not having the tool work at all on mobile.
-
-To address this, the layout of the main user interface needed to be redesigned, but the overall stack of web-based libraries and components chosen was already well-suited to mobile use.
-This effort is currently underway.
-
-\section{Conclusions and outlook}
-
-Experimenting with the various technologies and getting all of the disparate pieces working together was an enormous task, one which took much longer than anticipated.
-However, those decisions are now behind us and the platform has become quite stable.
-
-A new visualization can now be generated by the researchers in a matter of days or weeks.
-The researchers are admittedly very familiar with the inner workings of the system, but even so it has been encouraging to see new visualizations go from ideation to rough draft in such a short time.
-
-None of the above-listed visualizations are particularly groundbreaking or visually stunning.
-And, all of them could be easily created in other tools (although usually without the interactivity that the web enables).
-This is a bit disappointing but the open nature of the platform, requiring no software installation by end users, still has an advantage: it opens up the display of MATSim results to the public and to decisionmakers, even if they do not have access to desktop mapping or travel forecasting software.
-
-Another use case that has emerged from these visualization experiments is a more regimented data management tool for MATSim.
-Currently there is no straightforward way to share MATSim datasets online.
-The combination of the new file storage and management capabilities along with the Github-like user interface provides a natural place for users to store and disseminate results.
-
-Importantly, the world has not stood still while this platform was under development.
-Just in the past year, major data visualization efforts from well-funded companies such as Uber and others have been released.
-There are legitimate questions about how much of this work could be superceded by large, well-funded, professional coding teams.
-
-Despite these concerns, the MATSim visualization framework is operational and is now just beginning to be useful for researchers at the department of its creation.
-This bodes well for further development in the near future.
-
-All code is available on the MATSim Github site, at \url{github.com/matsim-org/viz}.
-
-\subsection{Author Contribution Statement}
-
-The authors confirm contribution to the paper as follows: study conception and
-design: Billy Charlton, Janek Laudan; data collection: Billy Charlton; analysis and
-interpretation of results: Billy Charlton, Janek Laudan; draft manuscript
-preparation: Billy Charlton. All authors reviewed the results and approved
-the final version of the manuscript.
-
-% -------------------------------------------------------------
-
-\begin{acks}
-This research project is part of the National Research Programme "Big Data" (NRP 75) of the Swiss National Science Foundation.\\
-Website: \url{http://www.nrp75.ch/}
-\end{acks}
-
-% \bibliography{paper}
-%% \bibliographystyle{SageV}
-
-%-----------------------------
-% \iffalse
-\begin{thebibliography}{1}
-
-\bibitem{R1}
-Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016). \textit{The multi-agent transport simulation MATSim} (p. 618). London: Ubiquity Press.
-
-\bibitem{R2}
-Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}.
+%   \bibitem{R2}
+%   Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}.
 In Andreas Horni, Kai Nagel, Kay W.
@@
 
-\bibitem{R3}
-Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–-224.
+%   \bibitem{R3}
+%   Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–-224.
 
-\bibitem{R4}
-Free Software Soundation (2007) \textit{GNU General Public License}, Version 3.
-URL: www.gnu.org/licenses/gpl.html
+%   \bibitem{R4}
+%   Free Software Soundation (2007) \textit{GNU General Public License}, Version 3.
+%   URL: www.gnu.org/licenses/gpl.html
 
-\bibitem{R5}
-Satyanarayan, A., Moritz, D., Wongsuphasawat, K., and Heer, J.
+%   \bibitem{R5}
+%   Satyanarayan, A., Moritz, D., Wongsuphasawat, K., and Heer, J.
 (2016) \textit{Vega-Lite: A Grammar of Interactive Graphics}.
@@
 
-\bibitem{R6}
-Wilkinson, Leland (2005) \textit{The Grammar of Graphics}, Second Edition. Springer Press, Chicago, USA. DOI: doi.org/10.1007/0-387-28695-0
+%   \bibitem{R6}
+%   Wilkinson, Leland (2005) \textit{The Grammar of Graphics}, Second Edition. Springer Press, Chicago, USA. DOI: doi.org/10.1007/0-387-28695-0
 
-\end{thebibliography}
-%\fi
-%----------------------------
+%   \end{thebibliography}
 

### 2020-12-30 src/paper.tex
1st draft

@@
 \usepackage[utf8]{inputenc}
+\usepackage{csquotes}
 
@@
 Augmenting or replacing fixed-route transit lines with automated, connected, shared taxi fleets may be a desirable alternative for transit agencies in less-densely developed areas.
-The economics of fixed-route transit are difficult to square with sparse development patterns [*], and the advent of private ride-hailing services adds the "shared taxi" concept to people's travel options.
+The economics of fixed-route transit are difficult to square with sparse development patterns, especially with the advent of private ride-hailing services which add the "shared taxi" concept to people's travel options \cite{Hough2018}.
 The same technologies used by private ride-hailing, however, can also be modified and applied to future transit services.
 
-In this study, a new dynamic-response shared taxi service, or "DRT", is examined as an alternative approach to serving the population of rural areas and less-dense portions of urban regions.
+In this study, known in German as AVÖV\footnote{Automasierter vernetzter öffentlicher Verkehr: Automated, networked public transit}, a new dynamic-response shared taxi service, or "DRT", is examined as an alternative approach to serving the population of rural areas and less-dense portions of urban regions.
 Dynamic-response transit is analyzed using the MATSim agent-based transport microsimulation model.
@@
 %% ----- EXISTING TOOLS AND RESEARCH -----
-\section{Motivation: The State of Existing Tools}
+\section{Motivation: Existing Tools and Research}
 \label{motivation}
 
-- Existing tools are incapable of visualizing the outputs of the DRT module of MATSim in an efficient manner
-- Authors of the DRT modules of MATSim expressed an unfulfilled need to debug the code
-- Previous research on web-based MATSim visualization provides a proof-of-concept and extensible open-source platform on which to build DRT-specific visualizations; an obvious extension!
-- Agent-based microsimulation is quite complex and the reams of data and extensive details are not appropriate for public outreach and stakeholder meetings.
-PDF reports and powerpoint slides are the norm, but there is an opportunity here to build something truly interactive, possibly an order of magnitude more useful.
+Agent-based microsimulation is quite complex and the reams of data and overwhelming details are not appropriate for public outreach and stakeholder meetings.
+PDF reports and PowerPoint slides are the norm in these situations, but there is an opportunity here to build something truly interactive, possibly an order of magnitude more useful.
 
-%% ----- REQUIREMENTS -----
-\section{Requirements}
-\label{requirements}
+Many existing tools can analyze agent-based transport simulation outputs.
+\textbf{OTFVis} is a venerable open-source tool designed to integrate tightly with MATSim \cite{Srippgen2015OTFVisInBook}, however it is not under active development for several years and has no online component nor DRT-specific capabilities.
 
-- Open source code so others may build upon it and extend it
-- Works in any modern browser. IE11 and small players, no. (pull from other doc)
-- Must somehow manage massive output datasets from MATSim runs. Browsers can't load 800MB files support massive datasets; requires some sort of output post-processing and storage of results
-- Visually compelling and easy to organize results for various scenarios
-- Public facing scenarios and private debugging areas
+The commercial product \textbf{Senozon Via} \cite{Rieser2015SenozonViaInBook} is very feature rich, actively maintained, and is currently used by our research team for myriad internal analysis tasks.
+It has no online web-based component, although it is very easy to create high-quality screen recordings of simulation results.
+The lack of interactive options for outreach meetings, as well as Via's focus on advanced analyst tasks, make it a poor fit for outreach despite its many other advantages.
 
-%% ----- EXPERIMENTS -----
+General geographic information system tools, such as \textbf{ArcGis} and its open source corollary \textbf{QGis}, are again very analyst-focused.
+Unlike QGis, ArcGis can create high quality online dashboards and interactive maps, and study team has used them in other settings.
+ArcGis does not have any MATSim-specific capabilities; the rich agent-based details would be entirely lost in the aggregations required to get DRT outputs onto an ArcGis map.
+ArcGis is also commercial software and was not considered due to the required financial outlay.
+
+Research by \citet{CharltonLaudan2020WebBasedVisualization} shows that a web-based tool specific to MATSim is now possible, given the latest advances in 3D technology embedded in modern web browsers.
+That tool, known as "MatHub", includes a web front-end (website) which runs in the user's browser.
+Through this site the user organizes model outputs, and defines aggregate visualizations for a few typical use cases.
+The tool also requires a large, complex back-end server component which handles file storage, user authentication, output post-processing, and realtime visualization support.
+That work is an obvious starting point for this research, but for reasons that will be described below, much of it was ultimately discarded.
+
+In summary, the existing tools are all either incapable of visualizing the disaggregate outputs of the DRT module of MATSim in an efficient manner, have no web-based online component, or are inappropriate for anyone except highly trained analysts.
+Thus the new platform described below is seen as a viable solution.
+
+%% ----- THE PLATFORM -----
 \section{The Visualization Platform}
@@
 
-Platform
+Despite two years of development of the previous client/server "MatHub" solution for other studies, many problems remained.
+The two largest were: (1) The massive size of MATSim outputs (a single run can be 800Mb or more) always preclude loading them into a user's web browser.
+This forces post-processing and/or aggregation to happen, which can eliminate many of the benefits of a fully disaggregate analysis.
+(2) The back-end server design required extensive effort whenever new visualization capabilities were desired.
+A strict database design and overwrought user/group management capabilities resulted in a rock-solid system that was extremely onerous to modify.
+It turns out that complex file and user management was unnecessary.\footnote{This invokes the programming mantra "YAGNI" -- You Aren't Gonna Need It: just build the simplest possible solution that works instead of over-engineering things.}
 
-- MatHub platform had extensive back-end server infrastructure (six server images for user authentication, file storage, post-processing, real-time animations). But it sucked
-- MatHub front-end used MapBox API, but size of datasets made all but the simplest of scenarios difficult or impossible to visualize within browser memory constraints.
+Trial and error eventually converges on a simplified system that eliminates all of the back-end server complexity by eliminating all of the back-end servers:
 
-- Covid-19 portal threw out everything except centralized file storage, and relied on the analyst to run standardized post-processing scripts and store those smaller datasets.
-First experiments with ThreeJS were successful but very time consuming to write
-- deck.gl proved far more performant than MapBox, also open source, and easier to write custom "shaders" using deck.gl examples
+\begin{itemize}
+  \item File storage is delegated to existing online storage, instead of creating a special visualization server.
+On site, we have an existing "Subversion" server \cite{Collins-SussmanEtc2008SubversionBook} which is quite adept at large-file storage and version management.
+All access to the server is via a very simple interface that can be easily extended to cloud storage, such as Amazon S3
+  \item As the site is intended to be publicly accessible, we discard all user authentication capabilities.
+If you have a web browser, you can view the results of any runs which are stored \emph{in the public area} of the Subversion server.
+Naturally some parts of file storage remain off-limits to the greater Internet
+  \item Analysts define visualizations by creating small configuration text files in the same folder as model outputs.
+These files include all of the details of a visualization --- usually just a few lines --- pointing to the input files and configuration settings.
+This eliminates all the hassle of writing code for user interface input dialog boxes.
+  \item Post-processing of outputs must be performed by the analyst as part of their modeling workflow, before uploading to Subversion, since browsers cannot ingest MATSim outputs in their entirety.
+\end{itemize}
 
-DRT animation
+The code is designed as a "Single Page Application" or "SPA" which is a modern way of building Javascript-based websites that behave similarly to desktop applications.
+Most large consumer websites such as Twitter and Facebook behave in a similar manner, thus this paradigm is easy for end users to understand.
 
-- Initially just the vehicle animations
-- Analysts requested color by occupancy
-- origin/destination "flyovers"
+Again after much trial and error for which there is no space in this brief paper, the platform is the combination of many different advanced web technologies:
+
+\begin{itemize}
+  \item The overall application framework is based on \textbf{VueJS}, a popular Javascript toolkit for building interactive websites
+  \item The agent-based visualizations use custom code ("shaders") written using the \textbf{Deck.gl} visualization library
+  \item Older visualizations borrowed from Charlton's previous research use the \textbf{MapBox GL} mapping library
+  \item Extensive optimization of memory usage for handling these large datasets with the brilliant \textbf{crossfilter} Javascript library, which manages lightning-fast filtering of million-rows datasets in multiple dimensions
+\end{itemize}
+
+The entire codebase is open source and is designed with a "plug-in" architecture so that new visualizations can be developed quickly while relying on common infrastructure for things like file access.
+
+To create a new visualization, the developer copies an existing template, modifies the Javascript code to take advantage of the Deck.gl or MapBox libraries as needed, and registers it with the main application.
+End users can then create visualizations by adding corresponding configuration files to a scenario's file server folder.
 
@@
 
-{Strippgen2016}
-Back-end consists only of file storage.
-Our research team has an existing "Subversion" server which handles large files efficiently and retains file version history.
-Other file storage solutions such as cloud-based AWS etc could also be implemented (but weren't) -- all file management goes through a single File API which could have multiple back-ends.
+Several unique visualizations depict details specific to the dynamic-response transit scenarios in this study.
+Standard MATSim outputs such as mode shares by time of day, simulation statistics, trip length summaries, and so on are also available in each scenario subfolder.
 
-Front end code is a "single page app" that can be hosted on any web server platform. We are using Github Pages because it is free and fast.
+These descriptions fit in this scientific paper but for best results it is easiest to view the site in situ, as it remains online and is better demonstrated than described.\footnote{Site is viewable at avoev-vsp.github.io}
 
-Web portal front-end has a simple text-based configuration file which lists the scenarios to be publicly viewable
-- Analyst runs agent-based model and post-processing scripts to create road/transit network and DRT animation files
-- etc
+\subsection{DRT Animation}
+
+A 24-hour agent-based interactive animation of DRT vehicles traversing the roadway network includes the following features: (1) DRT vehicles traversing the roadway links in real-time, sped up, or in reverse time, for analysis; (2) Vehicles color-coded by occupancy, so users can see the relative amount of ride-sharing occurring at different times of day; (3) Passenger origin/destinations depicted as "flyovers" instead of on the street network, so actual supply and demand can be viewed separately from the roads on which that demand occurs;
+
+Each of these components can be turned on or off to assist in reducing clutter in the animation. The 3D map can be zoomed or panned and its pitch angle adjusted to assist in viewing.
+
+[[PIC OF ANIMATION OR TWO ]]
+
+\subsection{DRT Passenger and Vehicle volumes}
+
+Aggregate summaries of roadway volumes depict total DRT pasenger or vehicle volumes in the entire simulation, or for specific times of day.
+The user can click on a specific link to see the diurnal (time of day) distribution of trips for that link, or can modify the time window being displayed for all links from total to a specific range, such as 6 A.M.
+to 9 A.M.
+if morning commute trips are being studied.
+
+Difference plots are also generated for some scenarios, especially comparing the DRT scenarios to the "base case" depicting current conditions.
+
+\subsection{Cross-Scenario Modal Shifts}
+
+Another aggregate visualization of total mode shift from one scenario to another, in this study comparing DRT scenarios to the base case.
+These visualizations really highlight the degree to which the new DRT mode is "stealing" trips from existing transit and bicycle modes versus from private vehicles.
 
@@
 
-- It works!
-- Website
-- Six public scenarios
-- Gregor's findings
-- Transit agency public outreach meetings
+Throughout summer 2020, our team iterated on dozens of DRT scenarios: varying pricing structures, presence of drivers, service coverage areas, maximum wait time goals, and so on.
+In August 2020 we began the first public meetings with transit agency staff, with the fully operational website at \url{https://avoev-vsp.github.io} (available in German only).
 
-FROM GREGOR
+The final site displays a front page listing the results for 6-8 scenarios in each region, covering a wide spectrum of DRT scenarios.
+Then for each scenario, visualizations include the individual agent animations (see above) with occupancy, origin/destination, and routing details; mode shift summaries vs.
+base case; vehicle volume, passenger volume, and volume differences on roadway links compared to base case; aggregate levels of tripmaking for trips originating or destined to geographical areas ("hexagon plots"); as well as detailed model output summary graphs depicting general taxi occupancy levels by time of day and more.
 
-Not sure whether this helps, but is what comes to my mind: The drt viz helped us understand better how the drt code works and find some issues, e.g.
+Each outreach meeting started with technical staff describing the study, the agent-based model, and the results for each alternative in overview; followed by small break-out groups online, in which people could explore the scenarios and discuss the findings amongs themselves; and finally a return en banc for group discussion.
+No problems with bugs or site performance were reported.
+
+Some direct feedback from analysts included:
+
+\begin{displayquote}\emph{
+  The drt viz helped us understand better how the drt code works and find some issues, e.g.
 when the system is congested weird things can happen and a large group of vehicles with one passenger per vehicle only moves from the same start to the same destination at the same time (in an attempt to save each passenger a few seconds it serves them separately and wastes vehicles which then are missing for the next requests).
 It is really nice to have requests and vehicle occupancy shown, Via did only the latter after some pre-processing.
-Other than that the website obviously made the avoev workshops easier.
+}\end{displayquote}
+
+\begin{displayquote}\emph{
+  The website obviously made the AVÖV workshops easier.
 Instead of only sharing our screen we could hand out a website where people could click themselves.
 Unfortunately it seemed that those listeners did not use the website a lot and rather kept listening to us.
-Some hypotheses: Maybe because sometimes a group of people shared one computer and their internet connection was bad, maybe because they are not used to that kind of interaction (maybe they always just listen to others' presentations unless they present themselve and discuss verbally later).
-For us the other plots like pax/link or veh/link are not entirely new, some used QGis to produce those.
-However maybe the website forced us to actually produce all those plots systematically instead of an "we could do that plot, but maybe it's not necessary, let's look into something else or raw data only to avoid the hassle of setting up QGis..."
+Some hypotheses: Maybe because sometimes a group of people shared one computer and their internet connection was bad, maybe because they are not used to that kind of interaction...
+}\end{displayquote}
+
+\begin{displayquote}\emph{
+  The other plots like passengers/link or vehicles/link are not entirely new; some used QGis to produce those.
+However, the website forced us to actually produce all those plots systematically instead of 'we could do that plot, but maybe it's not necessary, let's look into something else or raw data only to avoid the hassle of setting up QGis...'
+}\end{displayquote}
+
+The feedback that some of these visualizations are not entirely "new" is of course expected: tools such as QGis and Via do indeed provide some very nice visualization capabilities already.
+They are not online tools, however, and are not accessible to non-technical staff such as at a public outreach meeting.
+
+Given this level of feedback, we are pleased with the initial roll-out of the visualization platform, and expect to continue developing it for further analysis of dynamic-response transit systems.
 
@@
 
-This is the platform we've been trying to build for three years. It's efficient, extensible, and open source.
+Several years of trial and error led us to this specific combination of technologies and techniques.
+The research team is deeply indebted to the developers of so many components and libraries on which this is based, all of which are freely available online and without which this work would have been entirely impossible.
+The final product is more than parts merely cobbled together: a usable, useful tool for decision-making now exists, and the DRT-specific extensions serve as a template for other work.
 
-Further enhancements to the underlying platform and to the DRT visualization capabilities are expected.
+Paraphrasing Dennis Ritchie's early description of the UNIX system in \citet{Ritchie1978}, "Success lies not so much in new inventions but rather in the full exploitation of a carefully selected set of fertile ideas, and especially in showing that they can be keys to the implementation of a small yet powerful system."
 
-Plugin architecture means that other javascript developers can write plugins for their own agent-based models and their own use cases.
+Our successful experience using the platform in a public outreach setting confirms our belief that advanced agent-based simulations can be part of a decision framework and public policy discussions -- ultimately leading to better informed decision-making.
+
+Further enhancements to the underlying platform and to the DRT visualization capabilities are ongoing; the plugin architecture enables other Javascript developers to write plugins for their own agent-based models and their own use cases, if they so desire.
+
+The website remains online at \url{https://avoev-vsp.github.io}.
+All code is available at \url{https://github.com/avoev-vsp/avoev-vsp.github.io} and is licensed under the GNU GPL Version 3 \cite{FSF2007GnuGPL}.
+The authors welcome feedback and code contributions.
 
@@
 
-\bibliography{vsp,ref,book}
+\bibliography{bibliography}
 \bibliographystyle{elsarticle-harv}
-
-% \begin{thebibliography}{1}
-
-%   \bibitem{R1}
-%   Horni, A., Nagel, K., \& Axhausen, K. W. (Eds.). (2016). \textit{The multi-agent transport simulation MATSim} (p. 618). London: Ubiquity Press.
-
-%   \bibitem{R2}
-%   Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}.
-In Andreas Horni, Kai Nagel, Kay W.
-Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp.
-225-–234.
-
-%   \bibitem{R3}
-%   Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–-224.
-
-%   \bibitem{R4}
-%   Free Software Soundation (2007) \textit{GNU General Public License}, Version 3.
-%   URL: www.gnu.org/licenses/gpl.html
 

### 2020-12-30 src/paper.tex
first version w screenshots

@@
 
-[[PIC OF ANIMATION OR TWO ]]
+\begin{figure}[ht]
+\centering
+\begin{subfigure}[c]{0.4\textwidth}
+   \includegraphics[width=\linewidth]{images/fig-drt-vehicles.png}
+   \caption{DRT vehicles and routes}
+   \label{fig:vehicles}
+\end{subfigure}
+\begin{subfigure}[c]{0.4\textwidth}
+   \includegraphics[width=\linewidth]{images/fig-drt-flyovers.png}
+   \caption{DRT passenger origins and destinations}
+   \label{fig:flyover}
+\end{subfigure}
+\caption{Dynamic Response Transit (DRT) animation screenshots}
+\end{figure}
 
@@
 
-Difference plots are also generated for some scenarios, especially comparing the DRT scenarios to the "base case" depicting current conditions.
+Difference plots compare the DRT scenarios to the "base case" depicting current conditions.
+
+\begin{figure}[ht]
+  \centering
+  \begin{subfigure}[c]{0.32\textwidth}
+     \includegraphics[width=\linewidth]{images/fig-link-vols.png}
+     \caption{DRT Vehicle volumes, daily aggregation }
+     \label{fig:vehicles}
+  \end{subfigure}
+  \begin{subfigure}[c]{0.32\textwidth}
+     \includegraphics[width=\linewidth]{images/fig-mode-share.png}
+     \caption{Mode shift, example DRT alternative vs. base case}
+     \label{fig:flyover}
+  \end{subfigure}
+  \caption{Dynamic Response Transit (DRT) further visualization examples}
+\end{figure}
 
@@
 
+\small{
+
 \begin{displayquote}\emph{
@@
 }\end{displayquote}
+
+}
 

### 2020-12-30 src/paper.tex
submitted version 30-Dec-2020

@@
 \usepackage[utf8]{inputenc}
-\usepackage{csquotes}
+% make sure quotation marks look correct
+\usepackage [english]{babel}
+\usepackage [autostyle, english = american]{csquotes}
+\MakeOuterQuote{"}
 
@@
 Augmenting or even replacing fixed-route transit lines with automated, connected, shared taxi fleets may be a desirable alternative in less-densely developed areas.
-The MATSim agent-based transport microsimulation model is used to study scenarios including status quo; dynamically-dispatched fleets with drivers; and fully autonomous fleets.
-This paper focuses on a data visualization portal which includes many interactive views, such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more.
+The MATSim agent-based transport microsimulation model is used to study scenarios including the status quo, dynamically-dispatched fleets with drivers, and fully autonomous fleets.
+This paper focuses on a data visualization portal which includes many interactive views such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more.
 The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g.
@@
 In addition, developers of the MATSim DRT extension found the visualizations particularly useful for debugging both the algorithms and the scenario definitions.
-All code is open source and, while this specific study has a very specific use case, the portal has an extensible design that could be modified for other purposes.
+The code is entirely open source and, while this specific study has a rather esoteric use case, the visualization platform has an extensible design that could be modified for other purposes.
 
@@
 \begin{keyword}
-shared taxis; data visualization; matsim; public outreach
+dynamic response transit; data visualization; agent based simulation; matsim; shared taxis
 \end{keyword}
@@
 
-The existing methodology behind using MATSim to simulate DRT is well-described by [*] and is not the focus of this paper.
+The existing methodology behind using MATSim to simulate DRT is well-described by Bischoff et al \cite{Bischoff2020,BischoffMaciejewskiNagel2017SharedTaxiIITSC} and is not the focus of this paper.
 Rather, in the course of the effort to develop, analyze and compare various DRT scenarios, a need arose to create some truly unique data visualizations that could assist in improving and displaying the DRT scenarios themselves.
 A web-based portal is the result of this research.
-The portal includes many interactive views, such as agent (taxi) movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more.
+The portal includes many interactive views, such as taxi movements color-coded by number of passengers and trip request origins and destinations, changes in roadway and passenger volumes compared to a base case, and more.
 The agent-based simulation covers a 24 hour simulation period; analysts can hone in on specific times of day to examine, e.g.
@@
 
-The tool is now in operation for several small cities and rural regions in Germany, and was successfully used as an outreach tool in online public meetings.
-Project partners at transit agencies are interested in understanding the detailed implications of the DRT simulations, and without a visual, interactive component, that would be extremely difficult to provide.
+The tool is in operation for several small cities and rural regions in Germany, and was successfully used as an outreach tool in public meetings online.
+Project partners at transit agencies are interested in understanding the implications of the DRT simulations, and without a visual component, that would be extremely difficult to provide.
 
@@
 
-Agent-based microsimulation is quite complex and the reams of data and overwhelming details are not appropriate for public outreach and stakeholder meetings.
-PDF reports and PowerPoint slides are the norm in these situations, but there is an opportunity here to build something truly interactive, possibly an order of magnitude more useful.
+Agent-based microsimulation is quite complex, producing reams of data and overwhelming details that are not appropriate for stakeholder meetings.
+PDF reports and PowerPoint slides are the norm in these situations, but building something truly interactive instead could be much more engaging and useful for decision-making.
 
 Many existing tools can analyze agent-based transport simulation outputs.
-\textbf{OTFVis} is a venerable open-source tool designed to integrate tightly with MATSim \cite{Srippgen2015OTFVisInBook}, however it is not under active development for several years and has no online component nor DRT-specific capabilities.
+\textbf{OTFVis} \cite{Srippgen2015OTFVisInBook} is a venerable open-source tool designed to integrate tightly with MATSim, however it is not under active development and has neither online features nor DRT-specific capabilities.
 
@@
 General geographic information system tools, such as \textbf{ArcGis} and its open source corollary \textbf{QGis}, are again very analyst-focused.
-Unlike QGis, ArcGis can create high quality online dashboards and interactive maps, and study team has used them in other settings.
+Unlike QGis, ArcGis can create high quality online dashboards and interactive maps, and the study team has used them in other settings.
 ArcGis does not have any MATSim-specific capabilities; the rich agent-based details would be entirely lost in the aggregations required to get DRT outputs onto an ArcGis map.
@@
 
-Research by \citet{CharltonLaudan2020WebBasedVisualization} shows that a web-based tool specific to MATSim is now possible, given the latest advances in 3D technology embedded in modern web browsers.
-That tool, known as "MatHub", includes a web front-end (website) which runs in the user's browser.
+Research by \citet{CharltonLaudan2020WebBasedVisualization} shows that a web-based tool specific to MATSim is now possible, given the latest advances in graphic technology embedded in modern web browsers.
+That tool includes a "front-end" website which runs in the user's browser.
 Through this site the user organizes model outputs, and defines aggregate visualizations for a few typical use cases.
-The tool also requires a large, complex back-end server component which handles file storage, user authentication, output post-processing, and realtime visualization support.
+The tool also requires a large, complex "back-end" server component which handles file storage, user authentication, output post-processing, and realtime visualization support.
 That work is an obvious starting point for this research, but for reasons that will be described below, much of it was ultimately discarded.
@@
 
-Despite two years of development of the previous client/server "MatHub" solution for other studies, many problems remained.
-The two largest were: (1) The massive size of MATSim outputs (a single run can be 800Mb or more) always preclude loading them into a user's web browser.
+Despite two years of development of the above-mentioned client/server solution for other studies, problems in the old system are persistent.
+The two largest: (1) The massive size of MATSim outputs always preclude loading them into a user's web browser; a single DRT run can be 800Mb or more.
 This forces post-processing and/or aggregation to happen, which can eliminate many of the benefits of a fully disaggregate analysis.
-(2) The back-end server design required extensive effort whenever new visualization capabilities were desired.
-A strict database design and overwrought user/group management capabilities resulted in a rock-solid system that was extremely onerous to modify.
-It turns out that complex file and user management was unnecessary.\footnote{This invokes the programming mantra "YAGNI" -- You Aren't Gonna Need It: just build the simplest possible solution that works instead of over-engineering things.}
+(2) The back-end server design requires extensive effort whenever new visualization capabilities are desired.
+A strict database design and overwrought user/group management capabilities result in a rock-solid system that is extremely onerous to modify.
 
-Trial and error eventually converges on a simplified system that eliminates all of the back-end server complexity by eliminating all of the back-end servers:
+It turns out that complex file and user management is unnecessary.\footnote{This invokes the programming mantra "YAGNI" -- You Aren't Gonna Need It: build the simplest possible solution that works, instead of over-engineering things.} From trial and error and iteration emerges a simplified system that eliminates all of the back-end server complexity by eliminating all of the back-end servers:
 
 \begin{itemize}
-  \item File storage is delegated to existing online storage, instead of creating a special visualization server.
-On site, we have an existing "Subversion" server \cite{Collins-SussmanEtc2008SubversionBook} which is quite adept at large-file storage and version management.
-All access to the server is via a very simple interface that can be easily extended to cloud storage, such as Amazon S3
+  \item File storage is delegated to existing servers, instead of creating a special visualization file server.
+On site, we have an existing "Subversion" server \cite{Collins-SussmanEtc2008SubversionBook} which is adept at large-file storage and version history.
+All access to the server is via a simple interface that can be easily extended to cloud storage, such as Amazon S3
   \item As the site is intended to be publicly accessible, we discard all user authentication capabilities.
@@
 \begin{itemize}
-  \item The overall application framework is based on \textbf{VueJS}, a popular Javascript toolkit for building interactive websites
+  \item The overall application framework is based on \textbf{VueJS}, a popular Javascript toolkit for building interactive websites.
+VueJS is the glue between user interface clicks and the application code that runs underneath.
   \item The agent-based visualizations use custom code ("shaders") written using the \textbf{Deck.gl} visualization library
-  \item Older visualizations borrowed from Charlton's previous research use the \textbf{MapBox GL} mapping library
-  \item Extensive optimization of memory usage for handling these large datasets with the brilliant \textbf{crossfilter} Javascript library, which manages lightning-fast filtering of million-rows datasets in multiple dimensions
+  \item Older visualizations borrowed from Charlton et al's previous research use the \textbf{Mapbox GL} mapping library
+  \item Extensive optimization of memory usage for handling these large datasets with the brilliant \textbf{crossfilter} Javascript library, which manages lightning-fast filtering of millions-of-rows datasets in multiple dimensions
 \end{itemize}
@@
 The entire codebase is open source and is designed with a "plug-in" architecture so that new visualizations can be developed quickly while relying on common infrastructure for things like file access.
-
 To create a new visualization, the developer copies an existing template, modifies the Javascript code to take advantage of the Deck.gl or MapBox libraries as needed, and registers it with the main application.
@@
 
-These descriptions fit in this scientific paper but for best results it is easiest to view the site in situ, as it remains online and is better demonstrated than described.\footnote{Site is viewable at avoev-vsp.github.io}
+The descriptions below fit in the paper's length, but for best understanding of an interactive website, it is better to view the site in situ.
+The project site remains online and is better demonstrated than described.\footnote{Site is viewable at \href{https://avoev-vsp.github.io}{avoev-vsp.github.io}}
 
@@
 
-Another aggregate visualization of total mode shift from one scenario to another, in this study comparing DRT scenarios to the base case.
+Users can also \emph{compare} the aggregate total mode shift from one scenario to another, in this study comparing DRT scenarios to the base case.
 These visualizations really highlight the degree to which the new DRT mode is "stealing" trips from existing transit and bicycle modes versus from private vehicles.
@@
 
-Throughout summer 2020, our team iterated on dozens of DRT scenarios: varying pricing structures, presence of drivers, service coverage areas, maximum wait time goals, and so on.
-In August 2020 we began the first public meetings with transit agency staff, with the fully operational website at \url{https://avoev-vsp.github.io} (available in German only).
+Throughout summer 2020, our team iterated on dozens of DRT scenarios: varying pricing structures, presence of drivers vs.
+fully automated service, service coverage areas, maximum wait time goals, and so on.
+In August 2020 we began the first public meetings with transit agency staff, with the fully operational website at \href{https://avoev-vsp.github.io}{avoev-vsp.github.io} (available in German only).
 
@@
 The research team is deeply indebted to the developers of so many components and libraries on which this is based, all of which are freely available online and without which this work would have been entirely impossible.
-The final product is more than parts merely cobbled together: a usable, useful tool for decision-making now exists, and the DRT-specific extensions serve as a template for other work.
+The final product is more than parts merely cobbled together: a usable, useful tool for decision-making now exists, and the DRT-specific extensions serve as a template for further work.
 
@@
 
-The website remains online at \url{https://avoev-vsp.github.io}.
-All code is available at \url{https://github.com/avoev-vsp/avoev-vsp.github.io} and is licensed under the GNU GPL Version 3 \cite{FSF2007GnuGPL}.
+The website remains online at \href{https://avoev-vsp.github.io}{avoev-vsp.github.io}.
+All code is available at \href{https://github.com/avoev-vsp/avoev-vsp.github.io}{github.com/avoev-vsp} and is licensed under the GNU GPL Version 3 \cite{FSF2007GnuGPL}.
 The authors welcome feedback and code contributions.
